---
title: "VirMic_analysis.Rmd"
author: "CWigington"
output: pdf_document
---

World-wide locations from which samples were collected. Over 5,671 samples were collected by 25 cruises to create the dataset used in the analysis. The data includes only samples with both microbe and virus like particle samples from marine environments. Each black dot represents the location at which one or more samples were taken. Samples were taken in coastal, deep ocean, arctic, and tropical waters at both depths as shallow as 1 meter to up to 5,931 meters.

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
#check for packages needed to create the figures below
install.packages('https://cran.r-project.org/src/contrib/Archive/lmmfit/lmmfit_1.0.tar.gz', repos = NULL, type="source")

list.of.packages <- c("grid","rmarkdown", "ggplot2", "dplyr","maps","mapproj","lme4","smatr","scales","lmodel2","knitr","nlme","reshape2","lmmfit","gridExtra","plyr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, repos="http://cran.rstudio.com/")
#load pacakges
require(ggplot2)
require(plyr)
require(dplyr)
require(maps)
require(mapproj)
require(lme4)
require(smatr)
require(scales)
require(lmodel2)
require(knitr)
require(nlme)
require(reshape2)
require(lmmfit)
require(gridExtra)
require(rmarkdown)
require(grid)

dir.create("output", showWarnings = FALSE)
dir.create("output/figures", showWarnings = FALSE)
dir.create("output/figures/jpg", showWarnings = FALSE)
dir.create("output/figures/pdf", showWarnings = FALSE)
dir.create("output/data", showWarnings = FALSE)
data = read.csv("VIRBACDATA.txt")
#save(data, file = "output/data/DATA.RData")

calcRSQ <- function(obs,pred){
  ssTotal <- sum((obs-mean(obs))^2)
  ssResidual <- sum((obs - pred)^2)
  return(1-(ssResidual/ssTotal))
}

insert_minor <- function(major_labs, n_minor,blanks) {labs <- 
                              c( sapply( major_labs, function(x) c(x, rep("", blanks) ) ) )
                              labs[1:(length(labs)-n_minor)]}
#load("output/data/DATA.RData")
#write.table(data, file = "VirMic_data.csv", append = FALSE, sep = " ", row.names = F, col.names = TRUE, qmethod = c("escape", "double"))

```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/DATA.RData")

world <- map_data("world")
index <- which(world$long > 178 & world$long < 190.3)
clean.world <- world[-index, ]
data$DeepShallow <- factor(data$DEPTH > 100, labels=c('<=100m', '>100m'))
#summary(data$DeepShallow)
#>100m    <=100m
#2921      2750
#levels(data$Study)

worldmap <- ggplot(clean.world, aes(x=long, y=lat)) + 
  geom_polygon( aes(group=group), color='grey80',fill='grey80' )

worldmap<-worldmap + 
  theme(title = element_text(size=20),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"),           
        #text = element_text(size=32),aspect.ratio=1, panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())               

for (i in seq(1:nrow(data))){
  if (data$DEPTH[i] <= 100){
    data$DeepShallowColor[i] = "chartreuse3"
  }
  else {
    data$DeepShallowColor[i] = "blue"
  }
}

points<-as.data.frame(table(data$Study))
names(points) <- c("Study", "Freq")
points$logfreq<-ifelse(log10(points$Freq)>0, log10(points$Freq), points$logfreq<-0)
newrow <- data.frame(Study = c('Total'), Freq = c(nrow(data)), logfreq = c(log10(nrow(data))))
total<-rbind(points,newrow)
total1<-total[which(total$Freq > 0),]
#kable(total1[,1:2])
data<-merge(data,points,by="Study")

longs<-c()
lats<-c()
samples<-c()
points1<-as.matrix(table(data$long,data$lat))
for (long in seq(1,nrow(points1))){
  for (lat in seq(1,ncol(points1))){
    if (points1[long,lat] >0){
      #points[2,29]
      longs<-c(longs,rownames(points1)[long])
      lats<-c(lats,colnames(points1)[lat])
      samples<-c(samples,points1[long,lat])
    }
  }
}

counts<-as.data.frame(cbind(longs,lats,samples,log10(samples)+1))

names(counts) <- c("long", "lat","samples","log10samplesplusone")
data<-merge(data,counts,by=c("long","lat"))




data$DeepShallowsize<-ifelse(data$DeepShallow == '>100m',2,4)
deep_layer <- data[data$DeepShallow == '>100m',]
shallow_layer <- data[data$DeepShallow == '<=100m',]

map <- worldmap + 
  geom_point(data=data, aes(x=long, y=lat, col=data$DeepShallowColor)) +
  geom_point(data=deep_layer,aes(x=long, y=lat),shape=15, colour="chartreuse3", size=4, alpha=1) +
  geom_point(data=shallow_layer, aes(x=long, y=lat), shape=20, colour="blue", size=3, alpha=1) +
  scale_color_manual(name=element_blank(), values=c("chartreuse3","blue"), labels=c('>100m','<=100m')) +
  scale_shape_manual(name=element_blank(), values=c(15,20), labels=c('>100m','<=100m')) +
  scale_size_manual(name=element_blank(), values = c(0, 0), labels=c('>100m','<=100m')) +
  #scale_alpha_manual(guide='none', values = c(.1,1), labels=c('>100m','<=100m')) +
  #scale_alpha_manual(values = c(.1,1), labels=c('>100m','<=100m'), guide = guide_legend()) +
  xlab("Longitude") +
  ylab("Latitude") +
  theme(title = element_text(size=20),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"),
        #text = element_text(size=32),
        aspect.ratio=.5,
        legend.position = 'right',
        legend.key = element_rect(fill = "white"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.key.height = unit(.5,"line")) +
  guides(colour = guide_legend(override.aes = list(shape=c(15,20),size=c(7,7),color=c("blue","chartreuse3"),alpha=c(1,1))))
map

pdf("output/figures/pdf/map.pdf", width=10, height=6)
print(map)
dev.off()

jpeg("output/figures/jpg/map.jpg", width=500, height=300)
print(map)
dev.off()

save(total1, file = "output/data/TOTAL1.RData")
```


Bacterial abundance vs. viral abundance colored by depth category. Each point represents a biological sample and is colored according to its depth category, i.e. samples taken at depths of 100 meters or less are colored gree while samples taken at depths of greater than 100 meters are colored blue.


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/DATA.RData")
data$Study <- factor(data$Study)
data$log10VIRUS<-log10(data$VIRUS)
data$log10BACTERIA<-log10(data$BACTERIA)
data$Log10_VirBac_Ratio <- log10(data$VIRUS/data$BACTERIA)
insert_minor <- function(major_labs, n_minor,blanks) {labs <- 
                                                        c( sapply( major_labs, function(x) c(x, rep("", blanks) ) ) )
                                                      labs[1:(length(labs)-n_minor)]}
#nrow(data)
# 5,671
small <- data[which(!is.na(data$BACTERIA) & !is.na(data$VIRUS)), ]
small$DepthCategory[small$DEPTH <= 100] <- "<=100m"
small$DepthCategory[small$DEPTH >  100] <- ">100m"
small$col<-ifelse(small$DepthCategory == "<=100m","chartreuse3","blue")
small$pch<-ifelse(small$DepthCategory == "<=100m",17,16)
#nrow(small)
#5671

Allscatterplot<-ggplot(small, aes(x=log10BACTERIA, y=log10VIRUS, color=DepthCategory)) +
  geom_point(size=3,shape=20) +
  geom_density2d(colour="black", bins=10)+
  scale_color_manual(name="Depth", values=c("blue","chartreuse3")) +
  labs(list(x=expression('Microbes per mL (log'[10]*' scale)'), 
            y=expression('Viruses per mL (log'[10]*' scale)'))) + 
  theme(plot.title=element_text(size=32),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"),
        aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.key = element_rect(fill = "white"),
        legend.title = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.position = c(.2,.85)) +
  scale_x_continuous(breaks = c(seq(3.5,9,by=.25)), 
                     labels = c(insert_minor(seq(3.5,9,by=.5),1,1)), 
                     limits = c(3.5,7.5), 
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(seq(3.5,9,by=.25)), 
                     labels = c(insert_minor(seq(3.5,9,by=.5),1,1)), 
                     limits = c(4.75,9), 
                     expand = c(0,0))

Allscatterplot <- Allscatterplot + guides(colour = guide_legend(override.aes = list(shape=c(16,16),size=c(7,7))))

Allscatterplot
pdf("output/figures/pdf/Allscatterplot.pdf", width=6, height=6)
print(Allscatterplot)
dev.off()

jpeg("output/figures/jpg/Allscatterplot.jpg", width=500, height=500)
print(Allscatterplot)
dev.off()

save(small, file = "output/data/SMALL.RData")
```


Histograms of the ratio of viral abundance to Microbial abundance and the log of the ratio of viral abundance to Microbial abundance. The histograms of ratio of viral abundance to Microbial abundance is shown for depth categories. The distribution at the top of the panel represents samples taken at depths of 100 meters or less; samples taken at depths of greater than 100 meters are shown in the bottom panel. The red arrow at the top shows the position of the median value. The blue arrows shows the range of value between which 95 percent of the data is found. The domain of ratio values shown include 99\% of all of the ratio data points.

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALL.RData")

all<-quantile(small$Log10_VirBac_Ratio,probs=c(.025,.5,.975))
deep<-quantile(small$Log10_VirBac_Ratio[which(small$DepthCategory=='>100m')],probs=c(.025,.5,.975))
shallow<-quantile(small$Log10_VirBac_Ratio[which(small$DepthCategory=='<=100m')],probs=c(.025,.5,.975))
ann <- data.frame(
  iqr25 = c(shallow[1],deep[1]),
  iqr975=c(shallow[3],deep[3]),
  med=c(shallow[2],deep[2]),
  DepthCategory=c('<=100m','>100m'),
  medtext=c(round(10^shallow[2],2),round(10^deep[2],2)),
  text25=c(round(10^shallow[1],2),round(10^deep[1],2)),
  text975=c(round(10^shallow[3],2),round(10^deep[3],2))
)

h<-hist(small$Log10_VirBac_Ratio[which(small$DepthCategory=='<=100m')], breaks=c(seq(-1,max(small$Log10_VirBac_Ratio)+1,.1)), plot=F)
shallowmax<-max(h$counts)
h1<-hist(small$Log10_VirBac_Ratio[which(small$DepthCategory=='>100m')], breaks=c(seq(-1,max(log10(small$VIRUS/small$BACTERIA))+1,.1)), plot=F)
deepmax<-max(h1$counts)
#shallowmax<-334
#deepmax<-366

ann$medValue<-c(shallowmax,deepmax)
ann$medValuePlusOne<-c(shallowmax+1,deepmax+1)
ann$medValuePlusTen<-c(shallowmax+10,deepmax+10)
ann$medValuePlusTwenty<-c(shallowmax+20,deepmax+20)
ann$medValuePlusThirty<-c(shallowmax+30,deepmax+30)
require(grid)
logRatioDistribution<-ggplot(small, aes(x=Log10_VirBac_Ratio, y=..count..)) +
  geom_histogram(binwidth=.1) +
  facet_grid(DepthCategory ~ .) +
  geom_segment(data=ann,aes(x=med,xend=med,y=medValuePlusTwenty,yend=medValue),arrow=arrow(length=unit(0.5,"cm")),show_guide=F, color="#D55E00") +
  geom_segment(data=ann,aes(x=iqr25,xend=iqr25,y=-20,yend=0),arrow=arrow(length=unit(0.5,"cm")),show_guide=F, color="#0072B2") +
  geom_segment(data=ann,aes(x=iqr975,xend=iqr975,y=-20,yend=0),arrow=arrow(length=unit(0.5,"cm")),show_guide=F, color="#0072B2") +
  geom_text(data=ann,aes(x=med,y=medValuePlusThirty,label=medtext),size=5,show_guide=F) +
  geom_text(data=ann,aes(x=iqr25,y=-30,label=text25),size=5,show_guide=F) +
  geom_text(data=ann,aes(x=iqr975,y=-30,label=text975),size=5,show_guide=F) +
  scale_x_continuous(limits=c(min(na.omit(small$Log10_VirBac_Ratio)),max(na.omit(small$Log10_VirBac_Ratio)))) +
  labs(list(x=expression('Ratio of viral abundance to microbial abundance (on log'[10]*' scale)'), 
            y='Frequency'))+ 
  theme(plot.title=element_text(size=32),
        aspect.ratio=1,
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"),
        #text = element_text(size=32),aspect.ratio=1, panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())

logRatioDistribution

pdf("output/figures/pdf/logRatioDistribution.pdf", width=8, height=10)
print(logRatioDistribution)
dev.off()

jpeg("output/figures/jpg/logRatioDistribution.jpg", width=400, height=500)
print(logRatioDistribution)
dev.off()
```


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALL.RData")

all<-quantile(small$VIRUS/small$BACTERIA,probs=c(.025,.5,.975))
deep<-quantile((small$VIRUS/small$BACTERIA)[which(small$DepthCategory=='>100m')],probs=c(.025,.5,.975))
shallow<-quantile((small$VIRUS/small$BACTERIA)[which(small$DepthCategory=='<=100m')],probs=c(.025,.5,.975))
xmax<-quantile((small$VIRUS/small$BACTERIA)[which(small$DepthCategory=='<=100m')],probs=c(.99))

ann1 <- data.frame(
  iqr25 = c(shallow[1],deep[1]),
  iqr975=c(shallow[3],deep[3]),
  med=c(shallow[2],deep[2]),
  DepthCategory=c('<=100m','>100m'),
  medtext=c(round(shallow[2],2),round(deep[2],2)),
  text25=c(round(shallow[1],2),round(deep[1],2)),
  text975=c(round(shallow[3],2),round(deep[3],2))
)
small$VirBac_Ratio<-small$VIRUS/small$BACTERIA
h<-hist(small$VirBac_Ratio[which(small$DepthCategory=='<=100m')], breaks=c(seq(-1,max(small$VirBac_Ratio)+1)), plot=F)
shallowmax<-max(h$counts)
h1<-hist(small$VirBac_Ratio[which(small$DepthCategory=='>100m')], breaks=c(seq(-1,max(small$VirBac_Ratio))+1,.1), plot=F)
deepmax<-max(h1$counts)
#shallowmax<-30
#deepmax<-134
ann1$maxValue<-c(shallowmax,deepmax)
ann1$maxValuePlusTen<-c(shallowmax+10,deepmax+10)
ann1$maxValuePlusTwenty<-c(shallowmax+20,deepmax+20)
ann1$maxValuePlusThirty<-c(shallowmax+30,deepmax+30)

RatioDistribution<-ggplot(small, aes(x=VIRUS/BACTERIA, y=..count..)) +
  geom_histogram(binwidth=1) +
  facet_grid(DepthCategory ~ .) +
  #facet_grid(. ~ DepthCategory) +
  geom_segment(data=ann1,aes(x=med,xend=med,y=maxValuePlusTwenty,yend=ann1$maxValue),arrow=arrow(length=unit(0.5,"cm")),show_guide=F, color="#D55E00")+
  geom_segment(data=ann1,aes(x=iqr25,xend=iqr25,y=-20,yend=0),arrow=arrow(length=unit(0.5,"cm")),show_guide=F, color="#0072B2")+
  geom_segment(data=ann1,aes(x=iqr975,xend=iqr975,y=-20,yend=0),arrow=arrow(length=unit(0.5,"cm")),show_guide=F, color="#0072B2")+
  geom_text(data=ann1,aes(x=med+10,y=maxValuePlusThirty,label=medtext),size=5,show_guide=F) +
  geom_text(data=ann1,aes(x=iqr25,y=-30,label=text25),size=5,show_guide=F) +
  geom_text(data=ann1,aes(x=iqr975,y=-30,label=text975),size=5,show_guide=F) +
  #xlim(0,xmax)+
  scale_x_continuous(limits=c(0, xmax),breaks=NULL) +
  labs(list(x=expression('\nVirus abundance per mL to \nmicrobe abundance per mL ratio (log'[10]*')'), 
            y='Frequency'))+ 
  #title='Distribution of the ratio of viral \nto bacterial abundance by depth')) +
  #title='Distribution of the ratio of Virus \nto Bacteria abundance by depth')) +
  theme(plot.title=element_text(size=32),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"),
        #text = element_text(size=32),aspect.ratio=1, panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        aspect.ratio=1)

RatioDistribution

pdf("output/figures/pdf/RatioDistribution.pdf", width=8, height=10)
print(RatioDistribution)
dev.off()

jpeg("output/figures/jpg/RatioDistribution.jpg", width=400, height=500)
print(RatioDistribution)
dev.off()
```

The log (base 10) of microbe abundance versus the log (base 10) of viral abunndance. Each point represents a biological sample. The black line is a 1:1 line while the red line represents the regression line fit by minimizing the sum of squares for the equation Y = $\alpha$X + $\beta$. Here $\alpha$ is 0.422 and $\beta$ is 4.451.

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALL.RData")
index <- which( small$DepthCategory == "<=100m")
smallLabSet<-small[index, ]
m1<-lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA)
ms <- summary(m1)
RSQ10BacModel<-list(r2 = format(1-(sum((smallLabSet$log10VIRUS-(1+(1*smallLabSet$log10BACTERIA)))^2)/
                                     sum((smallLabSet$log10VIRUS-mean(smallLabSet$log10VIRUS))^2)), digits = 3))
l <- list(r2 = format(ms$r.squared, digits = 3))
lmeq <- substitute(italic('Power Law') == r2,l)
lmeqstr <- as.character(as.expression(lmeq))
Teneq <- substitute(italic('10:1line') == r2,RSQ10BacModel)
Teneqstr <- as.character(as.expression(Teneq))
sma<-sma(smallLabSet$log10VIRUS ~smallLabSet$log10BACTERIA, method="SMA")
sma1<-sma(smallLabSet$VIRUS ~smallLabSet$BACTERIA, method="SMA")
rsq<-c(round(sma$r2[[1]],3))
rsqdf<-data.frame(rsq)
smaeq <- substitute(italic('SMAreg') == rsq,rsqdf)
#smaeq <- substitute(italic(r)^2 == rsq,rsqdf)
smaeqstr <- as.character(as.expression(smaeq))
meaneq <- substitute(italic('Mean') == 0)
meaneqstr <- as.character(as.expression(meaneq))

shallowmods<-ggplot(smallLabSet, aes(x=log10BACTERIA, y=log10VIRUS)) +
  geom_point(size=1, col='steel blue') +
  geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=1) +
  geom_abline(intercept=1, slope=1, col='black', size=1) +
  #geom_density2d(colour="black", bins=10, alpha=.75) +
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 6.8, y = 7.1, label = 'Power Law', parse = F) + 
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 6.5, y = 7.9, label = '10:1', parse = F) + 
  geom_segment(aes(x = median(smallLabSet$log10BACTERIA), y = 4.8, xend = median(smallLabSet$log10BACTERIA), yend = 4.87), arrow = arrow(length = unit(0.5, "cm")), col='black') +
  geom_segment(aes(x = 3.5, y = median(smallLabSet$log10VIRUS)), xend = 3.57, yend = median(smallLabSet$log10VIRUS), arrow = arrow(length = unit(0.5, "cm")), col='black') +
  labs(list(#x=expression('microbes per mL (log'[10]*')'), 
    x=expression('Microbes per mL (log'[10]*')'), 
    y=expression('Viruses per mL (log'[10]*')')))+ 
  theme(aspect.ratio=1,
        plot.title=element_text(size=32),
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        #text = element_text(size=32),aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  scale_x_continuous(breaks = c(seq(2.5,8,by=.25)), 
                     labels = c(insert_minor(seq(2.5,8,by=.5),1,1)), 
                     limits = c(3.5,7.5), 
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(seq(4.5,9,by=.25)), 
                     labels = c(insert_minor(seq(4.5,9,by=.5),1,1)), 
                     limits = c(4.8,8.75), 
                     expand = c(0,0)) 
shallowmods

pdf("output/figures/pdf/shallowmods.pdf", width=8, height=10)
print(shallowmods)
dev.off()

jpeg("output/figures/jpg/shallowmods.jpg", width=400, height=500)
print(shallowmods)
dev.off()

save(smallLabSet, file = "output/data/SMALLLABSET.RData")
```


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALL.RData")
index <- which( small$DepthCategory == ">100m")
DeepSmall<-small[index, ]
m1<-lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA)
ms <- summary(m1)
RSQ10BacModel<-list(r2 = format(1-(sum((DeepSmall$log10VIRUS-(1+(1*DeepSmall$log10BACTERIA)))^2)/
                                     sum((DeepSmall$log10VIRUS-mean(DeepSmall$log10VIRUS))^2)), digits = 3))
l <- list(r2 = format(ms$r.squared, digits = 3))
lmeq <- substitute(italic('Power Law') == r2,l)
lmeqstr <- as.character(as.expression(lmeq))
Teneq <- substitute(italic('10:1line') == r2,RSQ10BacModel)
Teneqstr <- as.character(as.expression(Teneq))
sma<-sma(DeepSmall$log10VIRUS ~DeepSmall$log10BACTERIA, method="SMA")
sma1<-sma(DeepSmall$VIRUS ~DeepSmall$BACTERIA, method="SMA")
rsq<-c(round(sma$r2[[1]],3))
rsqdf<-data.frame(rsq)
smaeq <- substitute(italic('SMAreg') == rsq,rsqdf)
smaeqstr <- as.character(as.expression(smaeq))
meaneq <- substitute(italic('Mean') == 0)
meaneqstr <- as.character(as.expression(meaneq))

deepmods<-ggplot(DeepSmall, aes(x=log10BACTERIA, y=log10VIRUS)) +
  geom_point(size=1, col='steel blue') +
  geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=1) +
  geom_abline(intercept=1, slope=1, col='black', size=1) +
  #geom_density2d(colour="black", bins=10, alpha=.75) +
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 6.8, y = 7.1, label = 'Power Law', parse = F) + 
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 6.5, y = 7.9, label = '10:1', parse = F) + 
  geom_segment(aes(x = median(DeepSmall$log10BACTERIA), y = 4.8, xend = median(DeepSmall$log10BACTERIA), yend = 4.87), arrow = arrow(length = unit(0.5, "cm")), col='black') +
  geom_segment(aes(x = 3.5, y = median(DeepSmall$log10VIRUS)), xend = 3.57, yend = median(DeepSmall$log10VIRUS), arrow = arrow(length = unit(0.5, "cm")), col='black') +
  labs(list(#x=expression('Microbes per mL (log'[10]*')'), 
    x=expression('Microbes per mL (log'[10]*')'), 
    y=expression('Viruses per mL (log'[10]*')')))+ 
  theme(aspect.ratio=1,
        plot.title=element_text(size=32),
        axis.text=element_text(size=16), 
        axis.title=element_text(size=16,face="bold"),
        #text = element_text(size=32),aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) +
  scale_x_continuous(breaks = c(seq(2.5,8,by=.25)), 
                     labels = c(insert_minor(seq(2.5,8,by=.5),1,1)), 
                     limits = c(3.5,7.5), 
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(seq(4.5,9,by=.25)), 
                     labels = c(insert_minor(seq(4.5,9,by=.5),1,1)), 
                     limits = c(4.8,8.75), 
                     expand = c(0,0)) 
deepmods

pdf("output/figures/pdf/deepmods.pdf", width=8, height=10)
print(deepmods)
dev.off()

jpeg("output/figures/jpg/deepmods.jpg", width=400, height=500)
print(deepmods)
dev.off()

save(DeepSmall, file = "output/data/DEEPSMALL.RData")
```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData")
load("output/data/SMALL.RData")
load("output/data/DEEPSMALL.RData")
summary(small)
small$DeepShallow<-small$DepthCategory
ann4 <- data.frame(
  intset = c(summary(lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA))$coefficients[1],summary(lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA))$coefficients[1]),
  slopeset = c(summary(lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA))$coefficients[2],summary(lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA))$coefficients[2]),
  MEintset= c(summary(lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=smallLabSet))$coefficients$fixed[1],summary(lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=DeepSmall))$coefficients$fixed[1]),
  MEslopeset = c(summary(lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=smallLabSet))$coefficients$fixed[2],summary(lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=DeepSmall))$coefficients$fixed[2]),
  DeepShallow = c('<=100m', '>100m'),
  BacMedians = c(median(smallLabSet$log10BACTERIA), median(DeepSmall$log10BACTERIA)),
    VirMedians = c(median(smallLabSet$log10VIRUS), median(DeepSmall$log10VIRUS))
  )

ann5 <- data.frame(meX = 3.73, 
                   meY = 5.3, 
                   plX = 6.8,
                   plY = 6.9, 
                   toX=7, 
                   toY = 8.3, 
                   DeepShallow = factor('<=100m', levels = c('<=100m', '>100m')))

shallowDeepMods<-ggplot(small, aes(x=log10BACTERIA, y=log10VIRUS)) +
  geom_point(col='steel blue',pch=20, size=1) +
  facet_grid(. ~ DeepShallow) +
  geom_abline(data=ann4, aes(intercept=intset, slope=slopeset), col='#D55E00', size=1) + #orange power law line
  #geom_abline(data=ann4, aes(intercept=MEintset, slope=MEslopeset), col='#33CC33', size=1) + #green mixed effects Fixed Effects line
  geom_abline(intercept=1, slope=1, col='black', size=1) + #black line
  geom_segment(data=ann4,aes(x=BacMedians,xend=BacMedians,y=4.8,yend=4.9),arrow=arrow(length=unit(0.3,"cm")),show_guide=F, color="black") +
  geom_segment(data=ann4,aes(x=3.5,xend=3.6,y=VirMedians,yend=VirMedians),arrow=arrow(length=unit(0.3,"cm")),show_guide=F, color="black") +
   geom_text(data=ann5,aes(x=toX,y=toY,label='10:1'),show_guide=F,size=6,family="Helvetica", fontface="bold") + 
   geom_text(data=ann5,aes(x=plX,y=plY,label='Power Law'),show_guide=F,size=6,family="Helvetica", fontface="bold") + 
  labs(list(x=expression('Microbes per mL (log'[10]*' scale)'), 
            y=expression('Viruses per mL (log'[10]*' scale)')))+
  theme(panel.margin = unit(2, "lines"),
        strip.background = element_rect(fill="white"),
        strip.text.x = element_text(size = 20,face="bold"),
          aspect.ratio=1,
          plot.title=element_text(size=32),
          axis.text=element_text(size=12), 
          axis.title=element_text(size=16,face="bold"),
          #text = element_text(size=32),aspect.ratio=1,
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black")) +
    scale_x_continuous(breaks = c(seq(2.5,8,by=.25)), 
                     labels = c(insert_minor(seq(2.5,8,by=.5),1,1)), 
                     limits = c(3.5,7.5), 
                     expand = c(0,0)) +
    scale_y_continuous(breaks = c(seq(4.5,9,by=.25)), 
                     labels = c(insert_minor(seq(4.5,9,by=.5),1,1)), 
                     limits = c(4.8,8.75), 
                     expand = c(0,0))
shallowDeepMods
pdf("output/figures/pdf/shallowDeepMods.pdf", width=10, height=7)
print(shallowDeepMods)
dev.off()

jpeg("output/figures/jpg/shallowDeepMods.jpg", width=500, height=350)
print(shallowDeepMods)
dev.off()

```

Where the R-squared of the 10:1 model was calculated with log transformed data with an intercept of one and a coefficient of one, the r-squared values of a set of similar models were calculated to understand how varying the intercept by very small amounts relates to the r-squared of the model

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData")
m1<-lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA)
ms <- summary(m1)
ints<-c()
rsqs<-c()
for (int in seq(0,1.5,length=1000)){
  rsq<- 1-(sum((smallLabSet$log10VIRUS-(int+(smallLabSet$log10BACTERIA)))^2)/
             sum((smallLabSet$log10VIRUS-mean(smallLabSet$log10VIRUS))^2))
  rsqs<-c(rsqs,rsq)
  ints<-c(ints,int)
}

df<-as.data.frame(cbind(rsqs,ints))
nm<-as.character(expression('R'^{2}))
shallowlinearmods<-ggplot(data=df, aes(x=10^ints, y=rsqs)) +
  geom_point(size=1,col='steel blue')+
  geom_hline(yintercept=ms$r.squared) +
  geom_hline(yintercept=0, lty=2) +
  annotate(size=4,family="Helvetica", fontface="bold",geom="text",x = 10.25, y = 0.25, label = 'Power Law' , parse = F) +
  annotate(size=4,family="Helvetica", fontface="bold",geom="text",x = 30.25, y = -0.55, label = 'Fixed Ratio' , parse = F) +
  labs(list(x='Fixed VPR', 
            #         x=expression('10'^{alpha}), 
            y = expression(paste("Variance Explained (R"^"2",")"))))+
  theme(aspect.ratio=1,
        plot.title=element_text(size=20),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = c(.15,.8)) +
  scale_y_continuous(breaks = round(sort(c(seq(-5,1, length.out=7), ms$r.squared)),3),
                     #labels = c(insert_minor(seq(-6,2,by=1),0,0)), 
                     limits = c(-5,1.1), 
                     expand = c(0,0))
shallowlinearmods
pdf("output/figures/pdf/shallowlinearmods.pdf", width=8, height=10)
print(shallowlinearmods)
dev.off()

jpeg("output/figures/jpg/shallowlinearmods.jpg", width=400, height=500)
print(shallowlinearmods)
dev.off()
```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/DEEPSMALL.RData")
m1<-lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA)
ms <- summary(m1)
ints<-c()
rsqs<-c()
for (int in seq(0,1.5,length=1000)){
  rsq<- 1-(sum((DeepSmall$log10VIRUS-(int+(DeepSmall$log10BACTERIA)))^2)/
             sum((DeepSmall$log10VIRUS-mean(DeepSmall$log10VIRUS))^2))
  rsqs<-c(rsqs,rsq)
  ints<-c(ints,int)
}
df<-as.data.frame(cbind(rsqs,ints))
nm<-as.character(expression('R'^{2}))
deeplinearmods<-ggplot(data=df, aes(x=10^ints, y=rsqs)) +
  geom_point(size=1,col='steel blue')+
  geom_hline(yintercept=0, lty=2) +
  annotate(size=4,family="Helvetica", fontface="bold",geom="text",x = 10.25, y = 0.82, label = 'Power Law' , parse = F) +
  annotate(size=4,family="Helvetica", fontface="bold",geom="text",x = 30.25, y = -0.75, label = 'Fixed Ratio' , parse = F) +
  geom_hline(aes(yintercept=ms$r.squared)) +
  scale_y_continuous(breaks = round(sort(c(seq(-5,1, length.out=7), ms$r.squared)),3))+
  labs(list(x='Fixed VPR', 
            y = expression(paste("Variance Explained (R"^"2",")"))))+
  theme(aspect.ratio=1,
        plot.title=element_text(size=20),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"), 
        #text = element_text(size=32),aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = c(.15,.8))+
  scale_y_continuous(breaks = round(sort(c(seq(-5,1, length.out=7), ms$r.squared)),3),
                     #labels = c(insert_minor(seq(-6,2,by=1),0,0)), 
                     limits = c(-5,1.1), 
                     expand = c(0,0))
deeplinearmods
pdf("output/figures/pdf/deeplinearmods.pdf", width=8, height=10)
print(deeplinearmods)
dev.off()

jpeg("output/figures/jpg/deeplinearmods.jpg", width=400, height=500)
print(deeplinearmods)
dev.off()

```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData")
load("output/data/SMALL.RData")
load("output/data/DEEPSMALL.RData")

ann4 <- data.frame(
  zeroInt = c(0,0),
  powerlawInt = c(summary(lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA))$r.squared,summary(lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA))$r.squared),
  DeepShallow = c('<=100m', '>100m'),
  PLX = c(10,10), 
  PLY = c(.3,.78),
  TTOX = c(7,10),
  TTOY = c(-1,-1)
  )

ints<-c()
shallowrsqs<-c()
deeprsqs<-c()
for (int in seq(0,1.5,length=1000)){
  deeprsq<- 1-(sum((DeepSmall$log10VIRUS-(int+(DeepSmall$log10BACTERIA)))^2)/
             sum((DeepSmall$log10VIRUS-mean(DeepSmall$log10VIRUS))^2))
  shallowrsq<- 1-(sum((smallLabSet$log10VIRUS-(int+(smallLabSet$log10BACTERIA)))^2)/
             sum((smallLabSet$log10VIRUS-mean(smallLabSet$log10VIRUS))^2))
  shallowrsqs<-c(shallowrsqs,shallowrsq)
  deeprsqs<-c(deeprsqs,deeprsq)
  ints<-c(ints,int)
}
df<-data.frame(int = as.numeric(c(ints,ints)), rsq = as.numeric(c(shallowrsqs,deeprsqs)), DeepShallow = c(rep('<=100m',1000),rep('>100m',1000)))
df$int<-as.numeric(as.character(df$int))
df$rsq<-as.numeric(as.character(df$rsq))

shallowdeeplinearmods <- ggplot(data=df, aes(x=10^df$int, y=rsq)) +
  geom_point(size=1,col='steel blue') +
  facet_grid(. ~ DeepShallow) +
  geom_hline(yintercept=0, lty=2) +
  geom_hline(aes(yintercept = powerlawInt), data=ann4) +
  geom_text(data=ann4,aes(x=PLX,y=PLY,label='Power Law'),show_guide=F,size=6,family="Helvetica", fontface="bold") + 
  geom_text(data=ann4,aes(x=TTOX,y=TTOY,label='10:1'),show_guide=F,size=6,family="Helvetica", fontface="bold") + 
  labs(list(x='Fixed VMR', 
            y = expression(paste("Variance Explained (R"^"2",")"))))+
  theme(panel.margin = unit(2, "lines"),
        aspect.ratio=1,
        strip.background = element_rect(fill="white"),
        strip.text.x = element_text(size = 16,face="bold"),
        plot.title=element_text(size=20),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = c(.15,.8))+
  scale_y_continuous(breaks = round(sort(c(seq(-5,1, length.out=7), ann4$powerlawInt)),3),
                     #labels = c(insert_minor(seq(-6,2,by=1),0,0)), 
                     limits = c(-3,1.1), 
                     expand = c(0,0))
shallowdeeplinearmods
pdf("output/figures/pdf/shallowdeeplinearmods.pdf", width=12, height=8)
print(shallowdeeplinearmods)
dev.off()

jpeg("output/figures/jpg/shallowdeeplinearmods.jpg", width=600, height=400)
print(shallowdeeplinearmods)
dev.off()
```

Table of the number of samples taken by study.
```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData")
load("output/data/DEEPSMALL.RData")
shallowpoints<-as.data.frame(table(smallLabSet$Study))
deeppoints<-as.data.frame(table(DeepSmall$Study))
allpoints<-as.data.frame(merge(shallowpoints,deeppoints, by = 'Var1'))
allpoints$total<-allpoints$Freq.x+allpoints$Freq.y
names(allpoints) <- c("Study", "Shallow","Deep","Total")
newrow <- data.frame(Study = c('Total'), Shallow = c(sum(allpoints$Shallow)), Deep = c(sum(allpoints$Deep)), Total = c(sum(allpoints$Total)))
allpoints<-rbind(allpoints,newrow)
kable(allpoints[,1:4])
```

Table of percentiles for ensemble of slopes generated by bootstrapping the simple linear regression model
```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData") 

calcRSQ <- function(obs,pred){
  ssTotal <- sum((obs-mean(obs))^2)
  ssResidual <- sum((obs - pred)^2)
  return(1-(ssResidual/ssTotal))
}

reps<-100
powerlawRSQs<-c()
TenOneRSQs<-c()
slopes<-mat.or.vec(reps,1)
slopeavgs<-mat.or.vec(reps,1)
slopes1<-mat.or.vec(reps,1)
slopeavgs1<-mat.or.vec(reps,1)
slopes2<-mat.or.vec(length(levels(factor(smallLabSet$Study))),reps)
slopeavgs2<-mat.or.vec(length(levels(factor(smallLabSet$Study))),reps)
set.seed(1984)
pb <- txtProgressBar(min = 0, max = reps, style = 3)
for (i in seq(1:reps)){
  setTxtProgressBar(pb, i)
  bootstrapset<-data.frame()
  studies<-c()
  for (study in levels(factor(smallLabSet$Study))){
    studies<-c(studies,study)
    #print(lab)
    set<-smallLabSet[which(smallLabSet$Study==study),]
    #print(nrow(set))
    randomset <- set[sample(nrow(set), nrow(set), replace=T), ]
    bootstrapset<-rbind(bootstrapset,randomset)
  }
  lm<-lm(bootstrapset$log10VIRUS~bootstrapset$log10BACTERIA)
  clm<-coefficients(lm)
  clm<-clm[2]
  slopes[i]<-clm
  slopeavgs[i]<-mean(slopes[c(1:i)])
  
  me<-lmer(bootstrapset$log10VIRUS~bootstrapset$log10BACTERIA + (1|bootstrapset$Study),REML = F)
  cme<-coefficients(me)
  cme<-cme[[1]][1,2]
  slopes1[i]<-cme
  slopeavgs1[i]<-mean(slopes1[c(1:i)])
  
  me1<-lmer(bootstrapset$log10VIRUS~bootstrapset$log10BACTERIA + (bootstrapset$log10BACTERIA|bootstrapset$Study),REML = F)
  cme1<-coefficients(me1)
  cme1<-cme1[[1]]
  
  for (j in seq(1:nrow(cme1))){
    slopes2[j,i]<-cme1[j,2]
    slopeavgs2[j,i]<-mean(slopes2[j,c(1:i)])
  }
  
  powerlawRSQs<-c(powerlawRSQs,summary(lm)$r.squared)
  TenOneRSQs<-c(TenOneRSQs,calcRSQ(bootstrapset$log10VIRUS,bootstrapset$log10BACTERIA + 1))
  
}
close(pb)
percentiles<-as.matrix(quantile(slopes,probs=c(.025,.5,.975)))
colnames(percentiles) <- c("log10BACTERIA")

percentiles1<-as.matrix(quantile(slopes1,probs=c(.025,.5,.975)))
colnames(percentiles1) <- c("log10BACTERIA")

percentiles2<-mat.or.vec(length(levels(factor(smallLabSet$Study))),3)
for (j in seq(1:nrow(cme1))){
  quants<-as.matrix(quantile(slopes2[j,],probs=c(.025,.5,.975)))
  for (i in seq(1:3)){
    percentiles2[j,i]<-quants[i]
  }
}
percentiles2<-as.data.frame(percentiles2)
colnames(percentiles2) <- c("2.5%", "50%","97.5%")
rownames(percentiles2) <- c("ARCTICSBI","BATS","BEDFORDBASIN","NORTHSEA2001","CASES03-04","ELA","FECYCLE1","FECYCLE2","GEOTRACES","GEOTRACES_LEG3","KH04_5","KH05_2","GREENLAND2012","INDIANOCEAN2006","MOVE","NASB2005","POWOW","RAUNEFJORD2000","SOG","STRATIPHYT1","STRATIPHYT2","SWAT","TABASCO","TROUT","USC MO")

#save(powerlawRSQs, file = "output/data/POWERLAWRSQS.RData")
#save(TenOneRSQs, file = "output/data/TENONERSQS.RData")
#save(slopes2, file = "output/data/SLOPES2.RData")
#save(percentiles2, file = "output/data/SHALLOWPERCENTILES2.RData")

load("output/data/DEEPSMALL.RData")
slopes<-mat.or.vec(reps,1)
slopeavgs<-mat.or.vec(reps,1)
slopes1<-mat.or.vec(reps,1)
slopeavgs1<-mat.or.vec(reps,1)
slopes2<-mat.or.vec(length(levels(factor(DeepSmall$Study))),reps)
slopeavgs2<-mat.or.vec(length(levels(factor(DeepSmall$Study))),reps)
set.seed(1984)
pb <- txtProgressBar(min = 0, max = reps, style = 3)
for (i in seq(1:reps)){
  setTxtProgressBar(pb, i)
  bootstrapset<-data.frame()
  studies<-c()
  for (study in levels(factor(DeepSmall$Study))){
    studies<-c(studies,study)
    #print(lab)
    set<-DeepSmall[which(DeepSmall$Study==study),]
    #print(nrow(set))
    randomset <- set[sample(nrow(set), nrow(set), replace=T), ]
    bootstrapset<-rbind(bootstrapset,randomset)
  }
  lm<-lm(bootstrapset$log10VIRUS~bootstrapset$log10BACTERIA)
  clm<-coefficients(lm)
  clm<-clm[2]
  slopes[i]<-clm
  slopeavgs[i]<-mean(slopes[c(1:i)])
  
  me<-lmer(bootstrapset$log10VIRUS~bootstrapset$log10BACTERIA + (1|bootstrapset$Study),REML = F)
  cme<-coefficients(me)
  cme<-cme[[1]][1,2]
  slopes1[i]<-cme
  slopeavgs1[i]<-mean(slopes1[c(1:i)])
  
  me1<-lmer(bootstrapset$log10VIRUS~bootstrapset$log10BACTERIA + (bootstrapset$log10BACTERIA|bootstrapset$Study),REML = F)
  cme1<-coefficients(me1)
  cme1<-cme1[[1]]
  for (j in seq(1:nrow(cme1))){
    slopes2[j,i]<-cme1[j,2]
    slopeavgs2[j,i]<-mean(slopes2[j,c(1:i)])
  }
}
close(pb)
percentiles<-as.matrix(quantile(slopes,probs=c(.025,.5,.975)))
colnames(percentiles) <- c("log10BACTERIA")

percentiles1<-as.matrix(quantile(slopes1,probs=c(.025,.5,.975)))
colnames(percentiles1) <- c("log10BACTERIA")

percentiles2<-mat.or.vec(length(levels(factor(DeepSmall$Study))),3)
for (j in seq(1:nrow(cme1))){
  quants<-as.matrix(quantile(slopes2[j,],probs=c(.025,.5,.975)))
  for (i in seq(1:3)){
    percentiles2[j,i]<-quants[i]
  }
}
percentiles2<-as.data.frame(percentiles2)
colnames(percentiles2) <- c("2.5%", "50%","97.5%")
rownames(percentiles2) <- c("BATS","NORTHSEA2001","CASES03-04","GEOTRACES","GEOTRACES_LEG3","KH04_5","KH05_2","GREENLAND2012","INDIANOCEAN2006", "STRATIPHYT1","STRATIPHYT2","USC MO")
#kable(percentiles2)

deepslopes2<-slopes2
#save(deepslopes2, file = "output/data/DEEPSLOPES2.RData")
deeppercentiles2<-percentiles2
#save(deeppercentiles2, file = "output/data/DEEPPERCENTILES2.RData")

```

Create a scatterplot of all data with the study-specific scatterplots following
```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData")  
ARCTICSBI<-smallLabSet[which(smallLabSet$Study == 'ARCTICSBI'),]
FECYCLE2<-smallLabSet[which(smallLabSet$Study == 'FECYCLE2'),]
STRATAPHYT1<-smallLabSet[which(smallLabSet$Study == 'STRATIPHYT1'),]
MOVE<-smallLabSet[which(smallLabSet$Study == 'MOVE'),]
FECYCLE1<-smallLabSet[which(smallLabSet$Study == 'FECYCLE1'),]
USCMO<-smallLabSet[which(smallLabSet$Study == 'USC MO'),]
RAUNEFJORD2000<-smallLabSet[which(smallLabSet$Study == 'RAUNEFJORD2000'),]
KH05_2<-smallLabSet[which(smallLabSet$Study == 'KH05_2'),]
POWOW<-smallLabSet[which(smallLabSet$Study == 'POWOW'),]
GEOTRACES<-smallLabSet[which(smallLabSet$Study == 'GEOTRACES'),]
STRATAPHYT2<-smallLabSet[which(smallLabSet$Study == 'STRATIPHYT2'),]
SOG<-smallLabSet[which(smallLabSet$Study == 'SOG'),]
CASES0304<-smallLabSet[which(smallLabSet$Study == 'CASES03-04'),]
KH04_5<-smallLabSet[which(smallLabSet$Study == 'KH04_5'),]
GEOTRACES_LEG3<-smallLabSet[which(smallLabSet$Study == 'GEOTRACES_LEG3'),]
BEDFORDBASIN<-smallLabSet[which(smallLabSet$Study == 'BEDFORDBASIN'),]
SWAT<-smallLabSet[which(smallLabSet$Study == 'SWAT'),]
BATS<-smallLabSet[which(smallLabSet$Study == 'BATS'),]
GREENLAND2012<-smallLabSet[which(smallLabSet$Study == 'GREENLAND2012'),]
NORTHSEA2001<-smallLabSet[which(smallLabSet$Study == 'NORTHSEA2001'),]
INDIANOCEAN2006<-smallLabSet[which(smallLabSet$Study == 'INDIANOCEAN2006'),]
TABASCO<-smallLabSet[which(smallLabSet$Study == 'TABASCO'),]
TROUT<-smallLabSet[which(smallLabSet$Study == 'TROUT'),]
ELA<-smallLabSet[which(smallLabSet$Study == 'ELA'),]
NASB2005<-smallLabSet[which(smallLabSet$Study == 'NASB2005'),]
save(ARCTICSBI, file = "output/data/ARCTICSBI.RData")
save(FECYCLE2, file = "output/data/FECYCLE2.RData")
save(STRATAPHYT1, file = "output/data/STRATAPHYT1.RData")
save(MOVE, file = "output/data/MOVE.RData")
save(FECYCLE1, file = "output/data/FECYCLE1.RData")
save(USCMO, file = "output/data/USCMO.RData")
save(RAUNEFJORD2000, file = "output/data/RAUNEFJORD2000.RData")
save(KH05_2, file = "output/data/KH05_2.RData")
save(POWOW, file = "output/data/POWOW.RData")
save(GEOTRACES, file = "output/data/GEOTRACES.RData")
save(STRATAPHYT2, file = "output/data/STRATAPHYT2.RData")
save(SOG, file = "output/data/SOG.RData")
save(CASES0304, file = "output/data/CASES0304.RData")
save(KH04_5, file = "output/data/KH04_5.RData")
save(GEOTRACES_LEG3, file = "output/data/GEOTRACES_LEG3.RData") 
save(BEDFORDBASIN, file = "output/data/BEDFORDBASIN.RData")
save(SWAT, file = "output/data/SWAT.RData")
save(BATS, file = "output/data/BATS.RData")
save(GREENLAND2012, file = "output/data/GREENLAND2012.RData")
save(NORTHSEA2001, file = "output/data/NORTHSEA2001.RData")
save(INDIANOCEAN2006, file = "output/data/INDIANOCEAN2006.RData")
save(TABASCO, file = "output/data/TABASCO.RData")
save(TROUT, file = "output/data/TROUT.RData")
save(ELA, file = "output/data/ELA.RData")
save(NASB2005, file = "output/data/NASB2005.RData")
```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/ARCTICSBI.RData")
load("output/data/FECYCLE2.RData")
load("output/data/STRATAPHYT1.RData")
load("output/data/MOVE.RData")
load("output/data/FECYCLE1.RData")
load("output/data/USCMO.RData")
load("output/data/RAUNEFJORD2000.RData")
load("output/data/KH05_2.RData")
load("output/data/POWOW.RData")
load("output/data/GEOTRACES.RData")
load("output/data/STRATAPHYT2.RData")
load("output/data/SOG.RData")
load("output/data/CASES0304.RData")
load("output/data/KH04_5.RData")
load("output/data/GEOTRACES_LEG3.RData") 
load("output/data/BEDFORDBASIN.RData")
load("output/data/SWAT.RData")
load("output/data/BATS.RData")
load("output/data/GREENLAND2012.RData")
load("output/data/NORTHSEA2001.RData")
load("output/data/INDIANOCEAN2006.RData")
load("output/data/TABASCO.RData")
load("output/data/TROUT.RData")
load("output/data/ELA.RData")
load("output/data/NASB2005.RData")
load("output/data/SMALLLABSET.RData")
m1<-lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA)
fm2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=smallLabSet)
graphSmall<-function(dataset){
    main <- deparse(substitute(dataset))
    OBJ<-ggplot(dataset, aes(x=log10BACTERIA, y=log10VIRUS)) +
      geom_point(size=1, alpha = 1, col='steelblue') +
      labs(list(x='', y='', title=main))+
      geom_smooth(method='lm', se=F,size=.75,col='blue') +
      geom_abline(intercept=1, slope=1, col='black', size=.85, alpha=.3) +
      geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=.75, alpha=.5) +
      theme(aspect.ratio=1,
            title = element_text(size=8),
            axis.text=element_text(size=0),
            axis.title=element_text(size=0,face="bold"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.line = element_blank(),
            axis.ticks = element_blank(),
            legend.position = "NONE") +
      scale_x_continuous(#breaks = c(seq(4.5,9,by=.25)), 
        limits = c(min(smallLabSet$log10BACTERIA)-.1,max(smallLabSet$log10BACTERIA)+.1), 
        expand = c(0,0)) +
      scale_y_continuous(#breaks = c(seq(5.5,9,by=.25)), 
        limits = c(min(smallLabSet$log10VIRUS)-.1,max(smallLabSet$log10VIRUS)+.1), 
        expand = c(0,0))
    return(OBJ)
  }

graphSmall1<-function(dataset){
  main <- deparse(substitute(dataset))
  OBJ<-ggplot(dataset, aes(x=log10BACTERIA, y=log10VIRUS)) +
    geom_point(size=1, alpha = 1, col='steelblue') +
    labs(list(x='', y='', title=main))+
    geom_abline(intercept=1, slope=1, col='black', size=.85, alpha=.3) +
    geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=.75, alpha=.5) +
    geom_smooth(method='lm', se=F,size=.75,col='blue') +
    theme(aspect.ratio=1,
          title = element_text(size=8),
          axis.text.x=element_text(size=0),
          axis.text.y=element_text(size=6),
          axis.title=element_text(size=0,face="bold"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = c(.13,.75),
          legend.key = element_rect(fill = "white"),
          axis.line.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.key.size = unit(.5, "cm")) +
    scale_x_continuous(breaks = c(seq(4,9,by=1)), 
                       labels = c(insert_minor(seq(4,9,by=1),0,0)), 
                       limits = c(min(smallLabSet$log10BACTERIA)-.1,max(smallLabSet$log10BACTERIA)+.1), 
                       expand = c(0,0)) +
    scale_y_continuous(breaks = c(seq(5,9,by=1)), 
                       labels = c(insert_minor(seq(5,9,by=1),0,0)), 
                       limits = c(min(smallLabSet$log10VIRUS)-.1,max(smallLabSet$log10VIRUS)+.1), 
                       expand = c(0,0))
  return(OBJ)
}

graphSmall2<-function(dataset){
  main <- deparse(substitute(dataset))
  OBJ<-ggplot(dataset, aes(x=log10BACTERIA, y=log10VIRUS)) +
    geom_point(size=1, alpha = 1, col='steelblue') +
    labs(list(x='', y='', title=main))+
    geom_abline(intercept=1, slope=1, col='black', size=.85, alpha=.3) +
    geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=.75, alpha=.5) +
    geom_smooth(method='lm', se=F,size=.75,col='blue') +
    theme(aspect.ratio=1,
          title = element_text(size=8),
          axis.text.y=element_text(size=0),
          axis.text.x=element_text(size=6),
          axis.title=element_text(size=2,face="bold"),
          #text = element_text(size=32),aspect.ratio=1,
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = c(.13,.75),
          legend.key = element_rect(fill = "white"),
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.key.size = unit(.5, "cm")) +
    scale_x_continuous(breaks = c(seq(4,9,by=1)), 
                       labels = c(insert_minor(seq(4,9,by=1),0,0)), 
                       limits = c(min(smallLabSet$log10BACTERIA)-.1,max(smallLabSet$log10BACTERIA)+.1), 
                       expand = c(0,0)) +
    scale_y_continuous(breaks = c(seq(5,9,by=1)), 
                       labels = c(insert_minor(seq(5,9,by=1),0,0)), 
                       limits = c(min(smallLabSet$log10VIRUS)-.1,max(smallLabSet$log10VIRUS)+.1), 
                       expand = c(0,0))
  return(OBJ)
}

graphSmall3<-function(dataset){
  main <- deparse(substitute(dataset))
  OBJ<-ggplot(dataset, aes(x=log10BACTERIA, y=log10VIRUS)) +
    geom_point(size=1, alpha = 1, col='steelblue') +
    labs(list(x='', y='', title=main))+
    geom_abline(intercept=1, slope=1, col='black', size=.85, alpha=.3) +
    geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=.75, alpha=.5) +
    geom_smooth(method='lm', se=F,size=.75,col='blue') +
    theme(aspect.ratio=1,
          title = element_text(size=8),
          axis.text.y=element_text(size=6),
          axis.text.x=element_text(size=6),
          axis.title=element_text(size=2,face="bold"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = c(.13,.75),
          legend.key = element_rect(fill = "white"),
          legend.key.size = unit(.5, "cm")) +
    scale_x_continuous(breaks = c(seq(4,9,by=1)), 
                       labels = c(insert_minor(seq(4,9,by=1),0,0)), 
                       limits = c(min(smallLabSet$log10BACTERIA)-.1,max(smallLabSet$log10BACTERIA)+.1), 
                       expand = c(0,0)) +
    scale_y_continuous(breaks = c(seq(5,9,by=1)), 
                       labels = c(insert_minor(seq(5,9,by=1),0,0)), 
                       limits = c(min(smallLabSet$log10VIRUS)-.1,max(smallLabSet$log10VIRUS)+.1), 
                       expand = c(0,0))
  return(OBJ)
}



fig21a.1 <- ggplot(smallLabSet, aes(x=log10BACTERIA, y=log10VIRUS)) +
  geom_point(size=.01, alpha = .01) + 
  geom_line(aes(y=predict(fm2), group=Study), size=.25, alpha = 1, col="blue") +
  labs(list(x=expression('Microbes per mL (log'[10]*')'), 
            y=expression('Viruses per mL (log'[10]*')'))) +
  geom_abline(intercept=1, slope=1, col='black', size=2, alpha=.3) +
  geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=2, alpha=.5) +
  annotate(size=4,family="Helvetica", fontface="bold",geom="text",x = 6.75, y = 6.95, label = "Power Law", parse = F) + 
  annotate(size=4,family="Helvetica", fontface="bold",geom="text",x = 6.65, y = 7.95, label = as.character(paste("10","1",sep=":")), parse = F) +      
  theme(aspect.ratio=1,
        title = element_text(size=9),
        axis.text=element_text(size=7),
        axis.title=element_text(size=7,face="bold"),
        aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = 'NONE',
        legend.key = element_rect(fill = "white"),
        legend.key.size = unit(.5, "cm")) +
  scale_x_continuous(breaks = c(seq(4.5,9,by=.25)), 
                     labels = c(insert_minor(seq(4.5,9,by=.5),1,1)), 
                     limits = c(min(smallLabSet$log10BACTERIA)-.1,max(smallLabSet$log10BACTERIA)+.1), 
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(seq(5.5,9,by=.25)), 
                     labels = c(insert_minor(seq(5.5,9,by=.5),1,1)), 
                     limits = c(min(smallLabSet$log10VIRUS)-.1,max(smallLabSet$log10VIRUS)+.1), 
                     expand = c(0,0)) 

WIDTH<-7.5
HEIGHT<-9
vp1 <- viewport(width = 4.2/10, height = 4.3/12, x = 2/10, y = 10.1/12)
vp2 <- viewport(width = 3/10, height = 2.4 /12, x = 5/10, y = 11/12)
vp3 <- viewport(width = 3/10, height = 2.4/12, x = 7/10, y = 11/12)
vp4 <- viewport(width = 3/10, height = 2.4/12, x = 9/10, y = 11/12)
vp5 <- viewport(width = 3/10, height = 2.4/12, x = 5/10, y = 9/12)
vp6 <- viewport(width = 3/10, height = 2.4/12, x = 7/10, y = 9/12)
vp7 <- viewport(width = 3/10, height = 2.4/12, x = 9/10, y = 9/12)
vp8 <- viewport(width = 3/10, height = 2.4/12, x = 1/10, y = 7/12)
vp9 <- viewport(width = 3/10, height = 2.4/12, x = 3/10, y = 7/12)
vp10 <- viewport(width = 3/10, height = 2.4/12, x = 5/10, y = 7/12)
vp11 <- viewport(width = 3/10, height = 2.4/12, x = 7/10, y = 7/12)
vp12 <- viewport(width = 3/10, height = 2.4/12, x = 9/10, y = 7/12)
vp13 <- viewport(width = 3/10, height = 2.4/12, x = 1/10, y = 5/12)
vp14 <- viewport(width = 3/10, height = 2.4/12, x = 3/10, y = 5/12)
vp15 <- viewport(width = 3/10, height = 2.4/12, x = 5/10, y = 5/12)
vp16 <- viewport(width = 3/10, height = 2.4/12, x = 7/10, y = 5/12)
vp17 <- viewport(width = 3/10, height = 2.4/12, x = 9/10, y = 5/12)
vp18 <- viewport(width = 3/10, height = 2.4/12, x = 1/10, y = 3/12)
vp19 <- viewport(width = 3/10, height = 2.4/12, x = 3/10, y = 3/12)
vp20 <- viewport(width = 3/10, height = 2.4/12, x = 5/10, y = 3/12)
vp21 <- viewport(width = 3/10, height = 2.4/12, x = 7/10, y = 3/12)
vp22 <- viewport(width = 3/10, height = 2.4/12, x = 9/10, y = 3/12)
vp23 <- viewport(width = 3/10, height = 2.4/12, x = 1/10, y = 1/12)
vp24 <- viewport(width = 3/10, height = 2.4/12, x = 3/10, y = 1/12)
vp25 <- viewport(width = 3/10, height = 2.4/12, x = 5/10, y = 1/12)
vp26 <- viewport(width = 3/10, height = 2.4/12, x = 7/10, y = 1/12)
vp27 <- viewport(width = 3/10, height = 2.4/12, x = 9/10, y = 1/12)
grid.newpage()
pushViewport(viewport(layout = grid.layout(6,5)))

pdf("output/figures/pdf/scatterplots_by_study.pdf", width=8, height=8)
print(fig21a.1, vp = vp1)
print(graphSmall1(ARCTICSBI), vp = vp2)
print(graphSmall(BATS), vp = vp3)
print(graphSmall(BEDFORDBASIN), vp = vp4)
print(graphSmall1(CASES0304), vp = vp5)
print(graphSmall(ELA), vp = vp6)
print(graphSmall(FECYCLE1), vp = vp7)
print(graphSmall1(FECYCLE2), vp = vp8)
print(graphSmall(GEOTRACES), vp = vp9)
print(graphSmall(GEOTRACES_LEG3), vp = vp10)
print(graphSmall(GREENLAND2012), vp = vp11)
print(graphSmall(INDIANOCEAN2006), vp = vp12)
print(graphSmall1(KH04_5), vp = vp13)
print(graphSmall(KH05_2), vp = vp14)
print(graphSmall(MOVE), vp = vp15)
print(graphSmall(NASB2005), vp = vp16)
print(graphSmall(NORTHSEA2001), vp = vp17)
print(graphSmall1(POWOW), vp = vp18)
print(graphSmall(RAUNEFJORD2000), vp = vp19)
print(graphSmall(SOG), vp = vp20)
print(graphSmall(STRATAPHYT1), vp = vp21)
print(graphSmall2(STRATAPHYT2), vp = vp22)
print(graphSmall3(SWAT), vp = vp23)
print(graphSmall2(TABASCO), vp = vp24)
print(graphSmall2(TROUT), vp = vp25)
print(graphSmall2(USCMO), vp = vp26)
dev.off()

jpeg("output/figures/jpg/scatterplots_by_study.jpg", width=600, height=600)
print(fig21a.1, vp = vp1)
print(graphSmall1(ARCTICSBI), vp = vp2)
print(graphSmall(BATS), vp = vp3)
print(graphSmall(BEDFORDBASIN), vp = vp4)
print(graphSmall1(CASES0304), vp = vp5)
print(graphSmall(ELA), vp = vp6)
print(graphSmall(FECYCLE1), vp = vp7)
print(graphSmall1(FECYCLE2), vp = vp8)
print(graphSmall(GEOTRACES), vp = vp9)
print(graphSmall(GEOTRACES_LEG3), vp = vp10)
print(graphSmall(GREENLAND2012), vp = vp11)
print(graphSmall(INDIANOCEAN2006), vp = vp12)
print(graphSmall1(KH04_5), vp = vp13)
print(graphSmall(KH05_2), vp = vp14)
print(graphSmall(MOVE), vp = vp15)
print(graphSmall(NASB2005), vp = vp16)
print(graphSmall(NORTHSEA2001), vp = vp17)
print(graphSmall1(POWOW), vp = vp18)
print(graphSmall(RAUNEFJORD2000), vp = vp19)
print(graphSmall(SOG), vp = vp20)
print(graphSmall(STRATAPHYT1), vp = vp21)
print(graphSmall(STRATAPHYT2), vp = vp22)
print(graphSmall3(SWAT), vp = vp23)
print(graphSmall2(TABASCO), vp = vp24)
print(graphSmall2(TROUT), vp = vp25)
print(graphSmall2(USCMO), vp = vp26)
dev.off()
```


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/DEEPSMALL.RData")
m1<-lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA)
graphSmall<-function(dataset){
  main <- deparse(substitute(dataset))
  OBJ<-ggplot(dataset, aes(x=log10BACTERIA, y=log10VIRUS)) +
    geom_point(size=1, alpha = 1, col='steelblue') +
    labs(list(x='', y='', title=main))+
    geom_smooth(method='lm', se=F,size=.75,col='blue') +
    geom_abline(intercept=1, slope=1, col='black', size=.85, alpha=.3) +
    geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=.75, alpha=.5) +
    theme(aspect.ratio=1,
          title = element_text(size=8),
          axis.text=element_text(size=0),
          axis.title=element_text(size=0,face="bold"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "NONE") +
    scale_x_continuous(#breaks = c(seq(4.5,9,by=.25)),
      limits = c(min(DeepSmall$log10BACTERIA)-.1,max(DeepSmall$log10BACTERIA)+.1),
      expand = c(0,0)) +
    scale_y_continuous(#breaks = c(seq(5.5,9,by=.25)),
      limits = c(min(DeepSmall$log10VIRUS)-.1,max(DeepSmall$log10VIRUS)+.1),
      expand = c(0,0))
  return(OBJ)
}

graphSmall1<-function(dataset){
  main <- deparse(substitute(dataset))
  OBJ<-ggplot(dataset, aes(x=log10BACTERIA, y=log10VIRUS)) +
    geom_point(size=1, alpha = 1, col='steelblue') +
    labs(list(x='', y='', title=main))+
    geom_abline(intercept=1, slope=1, col='black', size=.85, alpha=.3) +
    geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=.75, alpha=.5) +
    geom_smooth(method='lm', se=F,size=.75,col='blue') +
    theme(aspect.ratio=1,
          title = element_text(size=8),
          axis.text.x=element_text(size=0),
          axis.text.y=element_text(size=6),
          axis.title=element_text(size=0,face="bold"),
          #text = element_text(size=32),aspect.ratio=1,
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = c(.13,.75),
          legend.key = element_rect(fill = "white"),
          axis.line.x = element_blank(),
          axis.ticks.x = element_blank(),
          legend.key.size = unit(.5, "cm")) +
    scale_x_continuous(breaks = c(seq(4,9,by=1)),
                       labels = c(insert_minor(seq(4,9,by=1),0,0)),
                       limits = c(min(DeepSmall$log10BACTERIA)-.1,max(DeepSmall$log10BACTERIA)+.1),
                       expand = c(0,0)) +
    scale_y_continuous(breaks = c(seq(5,9,by=1)),
                       labels = c(insert_minor(seq(5,9,by=1),0,0)),
                       limits = c(min(DeepSmall$log10VIRUS)-.1,max(DeepSmall$log10VIRUS)+.1),
                       expand = c(0,0))
  return(OBJ)
}

graphSmall2<-function(dataset){
  main <- deparse(substitute(dataset))
  OBJ<-ggplot(dataset, aes(x=log10BACTERIA, y=log10VIRUS)) +
    geom_point(size=1, alpha = 1, col='steelblue') +
    labs(list(x='', y='', title=main))+
    geom_abline(intercept=1, slope=1, col='black', size=.85, alpha=.3) +
    geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=.75, alpha=.5) +
    geom_smooth(method='lm', se=F,size=.75,col='blue') +
    theme(aspect.ratio=1,
          title = element_text(size=8),
          axis.text.y=element_text(size=0),
          axis.text.x=element_text(size=6),
          axis.title=element_text(size=2,face="bold"),
          #text = element_text(size=32),aspect.ratio=1,
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = c(.13,.75),
          legend.key = element_rect(fill = "white"),
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.key.size = unit(.5, "cm")) +
    scale_x_continuous(breaks = c(seq(4,9,by=1)),
                       labels = c(insert_minor(seq(4,9,by=1),0,0)),
                       limits = c(min(DeepSmall$log10BACTERIA)-.1,max(DeepSmall$log10BACTERIA)+.1),
                       expand = c(0,0)) +
    scale_y_continuous(breaks = c(seq(5,9,by=1)),
                       labels = c(insert_minor(seq(5,9,by=1),0,0)),
                       limits = c(min(DeepSmall$log10VIRUS)-.1,max(DeepSmall$log10VIRUS)+.1),
                       expand = c(0,0))
  return(OBJ)
}

graphSmall3<-function(dataset){
  main <- deparse(substitute(dataset))
  OBJ<-ggplot(dataset, aes(x=log10BACTERIA, y=log10VIRUS)) +
    geom_point(size=1, alpha = 1, col='steelblue') +
    labs(list(x='', y='', title=main))+
    geom_abline(intercept=1, slope=1, col='black', size=.85, alpha=.3) +
    geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=.75, alpha=.5) +
    geom_smooth(method='lm', se=F,size=.75,col='blue') +
    theme(aspect.ratio=1,
          title = element_text(size=8),
          axis.text.y=element_text(size=6),
          axis.text.x=element_text(size=6),
          axis.title=element_text(size=2,face="bold"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black"),
          legend.position = c(.13,.75),
          legend.key = element_rect(fill = "white"),
          legend.key.size = unit(.5, "cm")) +
    scale_x_continuous(breaks = c(seq(4,9,by=1)),
                       labels = c(insert_minor(seq(4,9,by=1),0,0)),
                       limits = c(min(DeepSmall$log10BACTERIA)-.1,max(DeepSmall$log10BACTERIA)+.1),
                       expand = c(0,0)) +
    scale_y_continuous(breaks = c(seq(5,9,by=1)),
                       labels = c(insert_minor(seq(5,9,by=1),0,0)),
                       limits = c(min(DeepSmall$log10VIRUS)-.1,max(DeepSmall$log10VIRUS)+.1),
                       expand = c(0,0))
  return(OBJ)
}

fm2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=DeepSmall)

fig21a.1 <- ggplot(DeepSmall, aes(x=log10BACTERIA, y=log10VIRUS)) +
  geom_point(size=.01, alpha = .01) +
  geom_line(aes(y=predict(fm2), group=Study), size=.25, alpha = 1, col="blue") +
  labs(list(x=expression('Microbes per mL (log'[10]*')'),
            y=expression('Viruses per mL (log'[10]*')'))) +
  geom_abline(intercept=1, slope=1, col='black', size=2, alpha=.3) +
  geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=2, alpha=.5) +
  annotate(size=4,family="Helvetica", fontface="bold",geom="text",x = 4.1, y = 6.1, label = 'Power\nLaw', parse = F) +
  annotate(size=4,family="Helvetica", fontface="bold",geom="text",x = 5.7, y = 7.0, label = "10:1", parse = F) +
  theme(aspect.ratio=1,
        title = element_text(size=9),
        axis.text=element_text(size=7),
        axis.title=element_text(size=7,face="bold"),
        aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = 'NONE',
        legend.key = element_rect(fill = "white"),
        legend.key.size = unit(.5, "cm")) +
  scale_x_continuous(breaks = c(seq(3,9,by=.25)),
                     labels = c(insert_minor(seq(3,9,by=.5),1,1)),
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(seq(3,9,by=.25)),
                     labels = c(insert_minor(seq(3,9,by=.5),1,1)),
                     expand = c(0,0))

vp1 <- viewport(width = 4.2/8, height = 4.3/10, x = 2/10, y = 10.1/12)
vp2 <- viewport(width = 2/8, height = 2/10, x = 1/8, y = 5/10)
vp3 <- viewport(width = 2/8, height = 2/10, x = 3/8, y = 5/10)
vp4 <- viewport(width = 2/8, height = 2/10, x = 5/8, y = 5/10)
vp5 <- viewport(width = 2/8, height = 2/10, x = 7/8, y = 5/10)
vp6 <- viewport(width = 2/8, height = 2/10, x = 1/8, y = 3/10)
vp7 <- viewport(width = 2/8, height = 2/10, x = 3/8, y = 3/10)
vp8 <- viewport(width = 2/8, height = 2/10, x = 5/8, y = 3/10)
vp9 <- viewport(width = 2/8, height = 2/10, x = 7/8, y = 3/10)
vp10 <- viewport(width = 2/8, height = 2/10, x = 1/8, y = 1/10)
vp11 <- viewport(width = 2/8, height = 2/10, x = 3/8, y = 1/10)
vp12 <- viewport(width = 2/8, height = 2/10, x = 5/8, y = 1/10)
vp13 <- viewport(width = 2/8, height = 2/10, x = 7/8, y = 1/10)

load("output/data/BATS.RData")
load("output/data/CASES0304.RData")
load("output/data/GEOTRACES.RData")
load("output/data/GEOTRACES_LEG3.RData")
load("output/data/GREENLAND2012.RData")
load("output/data/INDIANOCEAN2006.RData")
load("output/data/KH04_5.RData")
load("output/data/KH05_2.RData")
load("output/data/NORTHSEA2001.RData")
load("output/data/STRATAPHYT1.RData")
load("output/data/STRATAPHYT2.RData")
load("output/data/USCMO.RData")

grid.newpage()
pushViewport(viewport(layout = grid.layout(6,5)))
pdf("output/figures/pdf/deepscatterplots_by_study.pdf", width=8, height=8)
print(fig21a.1, vp = vp1)
print(graphSmall1(BATS), vp = vp2)
print(graphSmall(CASES0304), vp = vp3)
print(graphSmall(GEOTRACES), vp = vp4)
print(graphSmall(GEOTRACES_LEG3), vp = vp5)
print(graphSmall1(GREENLAND2012), vp = vp6)
print(graphSmall(INDIANOCEAN2006), vp = vp7)
print(graphSmall(KH04_5), vp = vp8)
print(graphSmall(KH05_2), vp = vp9)
print(graphSmall3(NORTHSEA2001), vp = vp10)
print(graphSmall2(STRATAPHYT1), vp = vp11)
print(graphSmall2(STRATAPHYT2), vp = vp12)
print(graphSmall2(USCMO), vp = vp13)
dev.off()

jpeg("output/figures/jpg/deepscatterplots_by_study.jpg", width=600, height=600)
print(fig21a.1, vp = vp1)
print(graphSmall1(BATS), vp = vp2)
print(graphSmall(CASES0304), vp = vp3)
print(graphSmall(GEOTRACES), vp = vp4)
print(graphSmall(GEOTRACES_LEG3), vp = vp5)
print(graphSmall1(GREENLAND2012), vp = vp6)
print(graphSmall(INDIANOCEAN2006), vp = vp7)
print(graphSmall(KH04_5), vp = vp8)
print(graphSmall(KH05_2), vp = vp9)
print(graphSmall3(NORTHSEA2001), vp = vp10)
print(graphSmall2(STRATAPHYT1), vp = vp11)
print(graphSmall2(STRATAPHYT2), vp = vp12)
print(graphSmall2(USCMO), vp = vp13)
dev.off()

```


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData")
load("output/data/SLOPES2.RData")
load("output/data/TOTAL1.RData")
#load("output/data/RSQFITS.RDATA")

insert_minor <- function(major_labs, n_minor,blanks) {labs <- 
                              c( sapply( major_labs, function(x) c(x, rep("", blanks) ) ) )
                              labs[1:(length(labs)-n_minor)]}

lme2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=smallLabSet)
ls2<-summary(lme2)
studies<-levels(smallLabSet$Study)
rsqs<-c()
studycol<-c()
pvals<-c()
for (i in seq(1:length(levels(smallLabSet$Study)))){
  sub<-smallLabSet[which(smallLabSet$Study == studies[i]),]
  foo <- sub[, c('log10VIRUS','log10BACTERIA')] 
  mod<-lm(sub$log10VIRUS ~ sub$log10BACTERIA)
  foo$yhat <- (sub$log10BACTERIA*coef(mod)[2])+coef(mod)[1]
  foo$e <- foo$log10VIRUS - foo$yhat
  SST <- sum( (foo$log10VIRUS - mean(foo$log10VIRUS))^2 )
  SSC <- sum( (foo$e)^2 )
  (SST-SSC)/SST
  studycol<-c(studycol,studies[i])
  rsqs<-c(rsqs,round((SST-SSC)/SST,3))
  summary<-summary(mod)
  pval<-summary$coefficients[2,4]
  pvals<-c(pvals,as.numeric(format(pval, scientific = TRUE, digits = 3)))
}
rsqfits<-as.data.frame(cbind(Study=studycol,RSQ=rsqs,'P-Value'=pvals),rownames=F)


fm2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=smallLabSet)
slopes<-as.data.frame(cbind(study=levels(factor(smallLabSet$Study)),slope=as.vector(coef(fm2)[,2])))
slopes$study <- factor(slopes$study, levels = slopes$study[order(slopes$slope)])
smallLabSet$Study <- factor(smallLabSet$Study, levels = rev(slopes$study[order(slopes$slope)]))

slopes2a<-as.data.frame(cbind(sort(levels(factor(smallLabSet$Study))),slopes2))
i <- sapply(slopes2a, is.factor)
slopes2a[i] <- lapply(slopes2a[i], as.character)
slopes2b<-melt(slopes2a, id.vars=c("V1"))
slopes2b<-as.data.frame(slopes2b)
x<-as.data.frame(cbind(as.character(slopes2b$V1),as.numeric(as.character(slopes2b$value))))
x$Study<-factor(x$V1)
x$slopes<-as.numeric(as.character(x$V2))
oind <- order(as.numeric(by(x$slopes, x$Study, median)))   
x$Study <- ordered(x$Study, levels=levels(x$Study)[oind])  

total2 <- total1[rev(order(total1$Freq)),]

smallLabSet1<-smallLabSet
smallLabSet1$Study <- factor(smallLabSet1$Study, levels = c("POWOW","TABASCO","FECYCLE2","FECYCLE1","NASB2005","SWAT","INDIANOCEAN2006","TROUT","STRATIPHYT2","SOG","GEOTRACES_LEG3","GREENLAND2012","MOVE","ELA","STRATIPHYT1","RAUNEFJORD2000","KH05_2","GEOTRACES","KH04_5","NORTHSEA2001","USC MO","BEDFORDBASIN","CASES03-04","ARCTICSBI","BATS"))

x1<-x
x1$Study <- factor(x$Study, levels = c("POWOW","TABASCO","FECYCLE2","FECYCLE1","NASB2005","SWAT","INDIANOCEAN2006","TROUT","STRATIPHYT2","SOG","GEOTRACES_LEG3","GREENLAND2012","MOVE","ELA","STRATIPHYT1","RAUNEFJORD2000","KH05_2","GEOTRACES","KH04_5","NORTHSEA2001","USC MO","BEDFORDBASIN","CASES03-04","ARCTICSBI","BATS"))

rsqtable<-rsqfits[order(as.numeric(format(pvals, scientific = TRUE, digits = 3))),]
rsqtable$labelcolor<-ifelse(as.numeric(levels(rsqtable$'P-Value'))[rsqtable$'P-Value']<.00192307, 'black','grey50')
x1<-merge(x1,rsqtable,by="Study")


levels(x1$Study)
p<-ggplot(x1, aes(x=Study, y=slopes)) +
  geom_violin(width=1.75, fill='lightblue') +
  geom_boxplot(width=.4,outlier.size = 0) +
  coord_flip() +
  geom_abline(intercept=1, slope=0, linetype=2) +
  ylab("Slope") +
  theme(title = element_text(size=20),
        axis.text=element_text(size=12, colour = "black"),
        #axis.text.y=element_text(size=12, colour = "black"),
        axis.title=element_text(size=16,face="bold"),
        axis.title.y=element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(colour = c("grey50", "grey50", "grey50", "grey50", "black", "grey50", "grey50", "black", "black",  "black",  "grey50",  "black", "black",  "black",  "black",  "black",  "black",  "black",  "black",  "black",  "black",  "black",  "black",  "black", "black",  "black"))) + 
  scale_y_continuous(breaks = c(seq(-.25,1.5,by=.25)), 
                     labels = c(insert_minor(seq(-.25,1.5,by=.25),0,0)), 
                     limits = c(0,1.5), 
                     expand = c(0,0))

levels(smallLabSet1$Study)
hist_right <- ggplot()+
  geom_histogram(aes(smallLabSet1$Study,fill=smallLabSet1$containsOneColor)) +
  scale_fill_manual(name="One in CI?", values = c("black","blue"), labels=c('No','Yes')) +
  coord_flip() +
  labs(list(y='Frequency')) + 
  theme(title = element_text(size=20),
        axis.text.x=element_text(size=10, colour="black"),
        axis.title=element_text(size=16,face="bold"),
        axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black"),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.key = element_rect(fill = "white"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 8),
        legend.key.height = unit(.2,"line"),
        legend.position = c(.4,.1)) + 
  scale_y_reverse()

WIDTH<-5
HEIGHT<-7
vp1 <- viewport(width = 8/10, height = 12/12, x = 4/10, y = 6/12)
vp2 <- viewport(width = 2/10, height = 12/12, x = 8.75/10, y = 6/12)

grid.newpage()
pushViewport(viewport(layout = grid.layout(6,5)))
print(hist_right, vp = vp2)
print(p, vp = vp1)

grid.newpage()
pushViewport(viewport(layout = grid.layout(6,5)))
pdf("output/figures/pdf/bootstrap_slopes.pdf", width=8, height=8)
print(hist_right, vp = vp2)
print(p, vp = vp1)
dev.off()

jpeg("output/figures/jpg/bootstrap_slopes.jpg", width=600, height=600)
print(hist_right, vp = vp2)
print(p, vp = vp1)
dev.off()

save(rsqfits, file='output/data/SHALLOWRSQFITS.RData')
```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/DEEPSMALL.RData")
load("output/data/DEEPSLOPES2.RData")
fm2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=DeepSmall)
slopes<-as.data.frame(cbind(study=levels(factor(DeepSmall$Study)),slope=as.vector(coef(fm2)[,2])))
slopes$study <- factor(slopes$study, levels = slopes$study[order(slopes$slope)])
DeepSmall$Study <- factor(DeepSmall$Study, levels = rev(slopes$study[order(slopes$slope)]))

slopes2a<-as.data.frame(cbind(sort(levels(factor(DeepSmall$Study))),deepslopes2))
#library(reshape2)
i <- sapply(slopes2a, is.factor)
slopes2a[i] <- lapply(slopes2a[i], as.character)
slopes2b<-melt(slopes2a, id.vars=c("V1"))
slopes2b<-as.data.frame(slopes2b)
x<-as.data.frame(cbind(as.character(slopes2b$V1),as.numeric(as.character(slopes2b$value))))
x$Study<-factor(x$V1)
x$slopes<-as.numeric(as.character(x$V2))
oind <- order(as.numeric(by(x$slopes, x$Study, median)))   
x$Study <- ordered(x$Study, levels=levels(x$Study)[oind])  

points<-as.data.frame(table(DeepSmall$Study))
names(points) <- c("Study", "Freq")
points$logfreq<-ifelse(log10(points$Freq)>0, log10(points$Freq), points$logfreq<-0)
newrow <- data.frame(Study = c('Total'), Freq = c(nrow(DeepSmall)), logfreq = c(log10(nrow(DeepSmall))))
total<-rbind(points,newrow)
DeepSmall<-merge(DeepSmall,points,by="Study")

DeepSmall1<-DeepSmall
DeepSmall1$Study <- factor(DeepSmall1$Study, levels = c("INDIANOCEAN2006","STRATIPHYT1","NORTHSEA2001","STRATIPHYT2","GREENLAND2012","CASES03-04","USC MO", "KH05_2", "GEOTRACES_LEG3", "KH04_5", "GEOTRACES","BATS"))

total <- total[order(total1$Freq[1:12]),] 
x1<-x
x1$Study <- factor(x$Study, levels = c("INDIANOCEAN2006","STRATIPHYT1","NORTHSEA2001","STRATIPHYT2","GREENLAND2012","CASES03-04","USC MO", "KH05_2", "GEOTRACES_LEG3", "KH04_5", "GEOTRACES","BATS"))

lme2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=DeepSmall)
ls2<-summary(lme2)
studies<-levels(DeepSmall$Study)
rsqs<-c()
studycol<-c()
pvals<-c()
for (i in seq(1:length(levels(DeepSmall$Study)))){
 sub<-DeepSmall[which(DeepSmall$Study == studies[i]),]
 foo <- sub[, c('log10VIRUS','log10BACTERIA')] 
 mod<-lm(sub$log10VIRUS ~ sub$log10BACTERIA)
 foo$yhat <- (sub$log10BACTERIA*coef(mod)[2])+coef(mod)[1]
 foo$e <- foo$log10VIRUS - foo$yhat
 SST <- sum( (foo$log10VIRUS - mean(foo$log10VIRUS))^2 )
 SSC <- sum( (foo$e)^2 )
 (SST-SSC)/SST
 studycol<-c(studycol,studies[i])
 rsqs<-c(rsqs,round((SST-SSC)/SST,3))
 summary<-summary(mod)
 pval<-summary$coefficients[2,4]
 pvals<-c(pvals,as.numeric(format(pval, scientific = TRUE, digits = 3)))
}
rsqfits<-as.data.frame(cbind(Study=studycol,RSQ=rsqs,'P-Value'=pvals),rownames=F)
kable(rsqfits)
rsqtable<-rsqfits[order(as.numeric(format(pvals, scientific = TRUE, digits = 3))),]
deeprsqfits<-rsqfits[order(as.numeric(format(pvals, scientific = TRUE, digits = 3))),]
rsqtable$labelcolor<-ifelse(as.numeric(levels(rsqtable$'P-Value'))[rsqtable$'P-Value']<.05/12, 'black','grey50')
x1<-merge(x1,rsqtable,by="Study")


p<-ggplot(x1, aes(x=Study, y=slopes)) +
 geom_violin(width=1.75, fill='lightblue') +
 geom_boxplot(width=.4,outlier.size = 0) +
 coord_flip() +
 geom_abline(intercept=1, slope=0, linetype=2) +
 ylab("Slope") +
 theme(title = element_text(size=20),
       axis.text=element_text(size=12, colour = "black"),
       #axis.text.y=element_text(size=12, colour = "black"),
       axis.title=element_text(size=16,face="bold"),
       axis.title.y=element_blank(),
       panel.background = element_blank(), 
       axis.line = element_line(colour = "black"),
       panel.grid.major = element_blank(),
       panel.grid.minor = element_blank(),
       axis.text.y = element_text(colour = c("grey50", "black", "black",  "black",  "black",  "grey50", "black",  "black",  "black",  "black",  "black",  "black"))) + 
 scale_y_continuous(breaks = c(seq(-.25,1.5,by=.25)), 
                    labels = c(insert_minor(seq(-.25,1.5,by=.25),0,0)), 
                    limits = c(0,1.5), 
                    expand = c(0,0))

hist_right <- ggplot()+
 #geom_histogram(aes(DeepSmall1$Study,fill=DeepSmall1$containsOneColor)) +
 geom_histogram(aes(DeepSmall1$Study,fill=DeepSmall1$containsOneColor)) +
 scale_fill_manual(name="One in CI?", values = c("black","blue"), labels=c('No','Yes')) +
 coord_flip() +
 labs(list(y='Frequency')) + 
 theme(title = element_text(size=20),
       axis.text.x=element_text(size=10, colour="black"),
       axis.title=element_text(size=16,face="bold"),
       axis.title.y=element_blank(),
       axis.text.y = element_blank(),
       panel.background = element_blank(), 
       axis.line = element_line(colour = "black"),
       axis.line.y = element_blank(),
       axis.ticks.y = element_blank(),
       panel.grid.major = element_blank(),
       panel.grid.minor = element_blank(),
       legend.key = element_rect(fill = "white"),
       legend.title = element_text(size = 12),
       legend.text = element_text(size = 8),
       legend.key.height = unit(.2,"line"),
       legend.position = c(.4,.1)) + 
scale_y_reverse()

WIDTH<-5
HEIGHT<-7
vp1 <- viewport(width = 8/10, height = 12/12, x = 4/10, y = 6/12)
vp2 <- viewport(width = 2/10, height = 12/12, x = 8.75/10, y = 6/12)
grid.newpage()
pushViewport(viewport(layout = grid.layout(6,5)))

pdf("output/figures/pdf/deepbootstrap_slopes.pdf", width=8, height=8)
print(hist_right, vp = vp2)
print(p, vp = vp1)
dev.off()

jpeg("output/figures/jpg/deepbootstrap_slopes.jpg", width=600, height=600)
print(hist_right, vp = vp2)
print(p, vp = vp1)
dev.off()

save(deeprsqfits, file='output/data/DEEPRSQFITS.RData')

```


Constrained regression model where the intercept for each lab was permitted to be between one standard deviation above or below the standard error of the intercept from the simple linear regression model. Here SE(intercept) = 0.10995505, estimate(intercept) = 4.4427386, and estimate(slope) = 0.4301043. R-squared for this model is 0.3486318.
```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData")
z4<-ifelse(smallLabSet$Study == "ARCTICSBI",1,0)
z24<-ifelse(smallLabSet$Study == "BATS",1,0)
z6<-ifelse(smallLabSet$Study == "BEDFORDBASIN",1,0)
z3<-ifelse(smallLabSet$Study == "NORTHSEA2001",1,0)
z8<-ifelse(smallLabSet$Study == "CASES03-04",1,0)
z12<-ifelse(smallLabSet$Study == "ELA",1,0)
z2<-ifelse(smallLabSet$Study == "FECYCLE1",1,0)
z22<-ifelse(smallLabSet$Study == "FECYCLE2",1,0)
z20<-ifelse(smallLabSet$Study == "GEOTRACES",1,0)
z19<-ifelse(smallLabSet$Study == "GEOTRACES_LEG3",1,0)
z16<-ifelse(smallLabSet$Study == "KH04_5",1,0)
z21<-ifelse(smallLabSet$Study == "KH05_2",1,0)
z5<-ifelse(smallLabSet$Study == "GREENLAND2012",1,0)
z25<-ifelse(smallLabSet$Study == "INDIANOCEAN2006",1,0)
z18<-ifelse(smallLabSet$Study == "MOVE",1,0)
z10<-ifelse(smallLabSet$Study == "NASB2005",1,0)
z23<-ifelse(smallLabSet$Study == "POWOW",1,0)
z17<-ifelse(smallLabSet$Study == "RAUNEFJORD2000",1,0)
z1<-ifelse(smallLabSet$Study == "SOG",1,0)
z9<-ifelse(smallLabSet$Study == "STRATIPHYT1",1,0)
z11<-ifelse(smallLabSet$Study == "STRATIPHYT2",1,0)
z26<-ifelse(smallLabSet$Study == "SWAT",1,0)
z15<-ifelse(smallLabSet$Study == "TABASCO",1,0)
z7<-ifelse(smallLabSet$Study == "TROUT",1,0)
z14<-ifelse(smallLabSet$Study == "USC MO",1,0)
y<-smallLabSet$log10VIRUS
x<-smallLabSet$log10BACTERIA
#summary(lm(log10VIRUS~log10BACTERIA,data=smallLabSet))$coef
#estimate + sd on lab intercepts
# 4.4427386 + 0.10995505 = 4.552694
# estimate - sd on lab intercepts
# 4.4427386 - 0.10995505 = 4.332784
variableInterceptwithconstraints<-nls(y~slope*x+SOG*z1+FECYCLE1*z2+NORTHSEA2001*z3+ARCTICSBI*z4+GREENLAND2012*z5+BEDFORDBASIN*z6+TROUT*z7+CASES0304*z8+STRATAPHYT1*z9+NASB2005*z10+STRATAPHYT2*z11+ELA*z12+USCMO*z14+TABASCO*z15+KHO4*z16+Raunefjord*z17+MOVE*z18+GEOTRACES_LEG3*z19+GEOTRACES*z20+KH05_2*z21+FECYCLE2*z22+POWOW*z23+BATS*z24+INDIANOCEAN2006*z25+SWAT*z26,algorithm="port",
start=c(slope=0,SOG=4.332784,FECYCLE1=4.332784,NORTHSEA2001=4.332784,ARCTICSBI=4.332784,GREENLAND2012=4.332784,BEDFORDBASIN=4.332784,TROUT=4.332784,CASES0304=4.332784,STRATAPHYT1=4.332784,NASB2005=4.332784,STRATAPHYT2=4.332784,ELA=4.332784,USCMO=4.332784,TABASCO=4.332784,KHO4=4.332784,Raunefjord=4.332784,MOVE=4.332784,GEOTRACES_LEG3=4.332784,GEOTRACES=4.332784,KH05_2=4.332784,FECYCLE2=4.332784,POWOW=4.332784,BATS=4.332784,INDIANOCEAN2006=4.332784,SWAT=4.332784),
lower=c(slope=-Inf,SOG=4.332784,FECYCLE1=4.332784,NORTHSEA2001=4.332784,ARCTICSBI=4.332784,GREENLAND2012=4.332784,BEDFORDBASIN=4.332784,TROUT=4.332784,CASES0304=4.332784,STRATAPHYT1=4.332784,NASB2005=4.332784,STRATAPHYT2=4.332784,ELA=4.332784,USCMO=4.332784,TABASCO=4.332784,KHO4=4.332784,Raunefjord=4.332784,MOVE=4.332784,GEOTRACES_LEG3=4.332784,GEOTRACES=4.332784,KH05_2=4.332784,FECYCLE2=4.332784,POWOW=4.332784,BATS=4.332784,INDIANOCEAN2006=4.332784,SWAT=4.332784),
upper=c(slope=Inf,SOG=4.552694,FECYCLE1=4.552694,NORTHSEA2001=4.552694,ARCTICSBI=4.552694,GREENLAND2012=4.552694,BEDFORDBASIN=4.552694,TROUT=4.552694,CASES0304=4.552694,STRATAPHYT1=4.552694,NASB2005=4.552694,STRATAPHYT2=4.552694,ELA=4.552694,USCMO=4.552694,TABASCO=4.552694,KHO4=4.552694,Raunefjord=4.552694,MOVE=4.552694,GEOTRACES_LEG3=4.552694,GEOTRACES=4.552694,KH05_2=4.552694,FECYCLE2=4.552694,POWOW=4.552694,BATS=4.552694,INDIANOCEAN2006=4.552694,SWAT=4.552694))
shallowcv<-variableInterceptwithconstraints
tss<-sum((smallLabSet$log10VIRUS-mean(smallLabSet$log10VIRUS))^2)
m1<-lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA)
shallowconstrained<-ggplot(smallLabSet, aes(x=log10BACTERIA, y=log10VIRUS, colour=Study)) +
  geom_point(size=.01, alpha=.01) +
  geom_abline(intercept=4.4806966, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.5526940, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.5526940, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3362559, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3399020, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.5526940, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3896576, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.5526940, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.5526940, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.5526940, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.332784, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.332784, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.5130411, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.5526940, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.4089481, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=4.3327840, slope=0.4301043, col='darkgreen', size=.25, alpha=1) +
  labs(list(x=expression(paste("Microbe abundance per mL (",log[10]," scale)")),
            y=expression(paste("Virus abundance per mL (",log[10]," scale)")))) +
  geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=2) +
  geom_abline(intercept=1, slope=1, col='#0072B2', size=2) +
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 5.0, y = 6.35, label = 'B', parse = TRUE) +
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 5.0, y = 6.85, label = 'A', parse = TRUE) +
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 6.75, y = 6.95, label = "Power Law", parse = F) +
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 6.65, y = 7.95, label = as.character(paste("10","1",sep=":")), parse = F) +
  theme(aspect.ratio=1,
        title = element_text(size=20),
        axis.text=element_text(size=16, colour="black"),
        axis.title=element_text(size=16,face="bold"),
        #text = element_text(size=32),aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        #legend.position = c(.1,.75),
        legend.position = 'none',
        legend.key = element_rect(fill = "white"),
        legend.key.size = unit(.5, "cm")) +
  scale_x_continuous(breaks = c(seq(4.5,9,by=.25)),
                     labels = c(insert_minor(seq(4.5,9,by=.5),1,1)),
                     #limits = c(4.6,9),
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(seq(5.5,9,by=.25)),
                     labels = c(insert_minor(seq(5.5,9,by=.5),1,1)),
                     #limits = c(min(smallLabSet$log10VIRUS)-1,max(smallLabSet$log10VIRUS)),
                     #expand = c(0,0)) + colScale
                     expand = c(0,0))

pdf("output/figures/pdf/constrainedRegression.pdf", width=8, height=8)
print(shallowconstrained)
dev.off()

jpeg("output/figures/jpg/constrainedRegression.jpg", width=600, height=600)
print(shallowconstrained)
dev.off()

shallowvariableInterceptwithconstraints<-variableInterceptwithconstraints
saveRDS(predict(variableInterceptwithconstraints), file = "output/data/SHALLOWVARIABLEINTERCEPTWITHCONSTRAINTS.rds")
save(shallowvariableInterceptwithconstraints, file = "output/data/SHALLOWCONSTRAINTS.RData")
```



```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/DEEPSMALL.RData")
STRATAPHYT1<-DeepSmall[which(DeepSmall$Study == 'STRATIPHYT1'),]
STRATAPHYT2<-DeepSmall[which(DeepSmall$Study == 'STRATIPHYT2'),]
CASES0304<-DeepSmall[which(DeepSmall$Study == 'CASES03-04'),]
USCMO<-DeepSmall[which(DeepSmall$Study == 'USC MO'),]
BATS<-DeepSmall[which(DeepSmall$Study == 'BATS'),]
KH05_2<-DeepSmall[which(DeepSmall$Study == 'KH05_2'),]
GEOTRACES<-DeepSmall[which(DeepSmall$Study == 'GEOTRACES'),]
KH04_5<-DeepSmall[which(DeepSmall$Study == 'KH04_5'),]
GEOTRACES_LEG3<-DeepSmall[which(DeepSmall$Study == 'GEOTRACES_LEG3'),]
GREENLAND2012<-DeepSmall[which(DeepSmall$Study == 'GREENLAND2012'),]
INDIANOCEAN2006<-DeepSmall[which(DeepSmall$Study == 'INDIANOCEAN2006'),]
NORTHSEA2001<-DeepSmall[which(DeepSmall$Study == 'NORTHSEA2001'),]
z24<-ifelse(DeepSmall$Study == "BATS",1,0)
z3<-ifelse(DeepSmall$Study == "NORTHSEA2001",1,0)
z8<-ifelse(DeepSmall$Study == "CASES03-04",1,0)
z20<-ifelse(DeepSmall$Study == "GEOTRACES",1,0)
z19<-ifelse(DeepSmall$Study == "GEOTRACES_LEG3",1,0)
z16<-ifelse(DeepSmall$Study == "KH04_5",1,0)
z21<-ifelse(DeepSmall$Study == "KH05_2",1,0)
z5<-ifelse(DeepSmall$Study == "GREENLAND2012",1,0)
z25<-ifelse(DeepSmall$Study == "INDIANOCEAN2006",1,0)
z9<-ifelse(DeepSmall$Study == "STRATIPHYT1",1,0)
z11<-ifelse(DeepSmall$Study == "STRATIPHYT2",1,0)
z14<-ifelse(DeepSmall$Study == "USC MO",1,0)
y<-DeepSmall$log10VIRUS
x<-DeepSmall$log10BACTERIA
#y%*%x/(sqrt(x%*%x)*sqrt(y%*%y)) = 0.9973633 i.e. theta = 180*(acos(0.9973633)/pi) = 4.16163 degrees
#summary(lm(log10VIRUS~log10BACTERIA,data=DeepSmall))$coef
#estimate + sd on lab intercepts
# 3.513389 + 0.037668749 = 3.551058
# estimate - sd on lab intercepts
# 3.513389 - 0.037668749 = 3.47572

#data$Study<-revalue(data$Study, c("Balsom - arctic"="ARCTICSBI",'Bratbak 2001'='NORTHSEA2001', "MiddleboeGREENLAND2012"="GREENLAND2012", "MiddleBoeIndia" = 'INDIANOCEAN2006'))

variableInterceptwithconstraints<-nls(y~slope*x+NORTHSEA2001*z3+GREENLAND2012*z5+CASES0304*z8+STRATAPHYT1*z9+STRATAPHYT2*z11+USCMO*z14+KHO4*z16+GEOTRACES_LEG3*z19+GEOTRACES*z20+KH05_2*z21+BATS*z24+INDIANOCEAN2006*z25,algorithm="port",
start=c(slope=0,NORTHSEA2001=3.47572,GREENLAND2012=3.47572,CASES0304=3.47572,STRATAPHYT1=3.47572,STRATAPHYT2=3.47572,USCMO=3.47572,KHO4=3.47572,GEOTRACES_LEG3=3.47572,GEOTRACES=3.47572,KH05_2=3.47572,BATS=3.47572,INDIANOCEAN2006=3.47572),
lower=c(slope=-Inf,NORTHSEA2001=3.47572,GREENLAND2012=3.47572,CASES0304=3.47572,STRATAPHYT1=3.47572,STRATAPHYT2=3.47572,USCMO=3.47572,KHO4=3.47572,GEOTRACES_LEG3=3.47572,GEOTRACES=3.47572,KH05_2=3.47572,BATS=3.47572,INDIANOCEAN2006=3.47572),
upper=c(slope=Inf,NORTHSEA2001=3.551058,GREENLAND2012=3.551058,CASES0304=3.551058,STRATAPHYT1=3.551058,STRATAPHYT2=3.551058,USCMO=3.551058,KHO4=3.551058,GEOTRACES_LEG3=3.551058,GEOTRACES=3.551058,KH05_2=3.551058,BATS=3.551058,INDIANOCEAN2006=3.551058))

tss<-sum((DeepSmall$log10VIRUS-mean(DeepSmall$log10VIRUS))^2)
deepcv<-variableInterceptwithconstraints
m1<-lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA)
deepconstrained<-ggplot(DeepSmall, aes(x=log10BACTERIA, y=log10VIRUS, colour=Study)) +
  geom_point(size=.01, alpha=.01) +
  geom_abline(intercept=3.4757200, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.4757200, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.4757200, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.5510580, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.5510580, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.5510580, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.5125687, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.4757200, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.5510580, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.5190784, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.4757200, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  geom_abline(intercept=3.4757200, slope=0.5325704, col='darkgreen', size=.25, alpha=1) +
  labs(list(x=expression(paste("Microbe abundance per mL (",log[10]," scale)")),
            y=expression(paste("Virus abundance per mL (",log[10]," scale)")))) +
  geom_abline(intercept=coef(m1)[[1]], slope=coef(m1)[[2]], col='#D55E00', size=2) +
  geom_abline(intercept=1, slope=1, col='#0072B2', size=2) +
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 4, y = 5.45, label = 'Power Law', parse = F) +
  annotate(size=6,family="Helvetica", fontface="bold",geom="text",x = 5.7, y = 7.0, label = "10:1", parse = F) +
  theme(aspect.ratio=1,
        plot.title=element_text(size=20),
        axis.text=element_text(size=16),
        axis.title=element_text(size=16,face="bold"),
        #text = element_text(size=32),aspect.ratio=1,
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        legend.position = "NONE") +
  scale_x_continuous(breaks = c(seq(3.0,8,by=.25)),
                     labels = c(insert_minor(seq(3.0,8,by=.5),1,1)),
                     #limits = c(4.6,8),
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(seq(3.0,9,by=.25)),
                     labels = c(insert_minor(seq(3.0,9,by=.5),1,1)),
                     #limits = c(4.6,9),
                     expand = c(0,0))

pdf("output/figures/pdf/deepconstrainedRegression.pdf", width=8, height=8)
print(deepconstrained)
dev.off()

jpeg("output/figures/jpg/deepconstrainedRegression.jpg", width=600, height=600)
print(deepconstrained)
dev.off()

deepfitteddata<-as.vector(predict(variableInterceptwithconstraints))
saveRDS(deepfitteddata, file = "output/data/DEEPVARIABLEINTERCEPTWITHCONSTRAINTS.rds")
deepvariableInterceptwithconstraints<-variableInterceptwithconstraints
save(deepvariableInterceptwithconstraints, file = "output/data/DEEPCONSTRAINTS.RData")
```


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
#figure22_1-1
require(grid)
load("output/data/SMALL.RData")
smallLabSet<-small[which(small$DepthCategory == "<=100m"),]
DeepSmall<-small[which(small$DepthCategory == ">100m"),]

ann4 <- data.frame(
  m1int = c(summary(lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA))$coefficients[1],summary(lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA))$coefficients[1]),
  m1slope = c(summary(lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA))$coefficients[2],summary(lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA))$coefficients[2]),
  DeepShallow = c('<=100m', '>100m'),
  constrainedslopes = c(0.4301043,0.5325704),
  constrainedints1 = c(4.4806966,3.4757200),
  constrainedints2 = c(4.5526940,3.5510580),
  constrainedints3 = c(4.3327840,3.5125687),
  constrainedints4 = c(4.3362559,3.5190784),
  constrainedints5 = c(4.3399020,3.5190784),
  constrainedints6 = c(4.3896576,3.5190784),
  constrainedints7 = c(4.5130411,3.5190784),
  constrainedints8 = c(4.4089481,3.5190784)
  )

ann5 <- data.frame(meX = 3.73, 
                   meY = 5.3, 
                   plX = 6.8,
                   plY = 6.9, 
                   toX=7, 
                   toY = 8.3,
                   Alab = 'A',
                   AX = 4.5, 
                   AY = 6.75,
                   BX = 4.5, 
                   BY = 6,
                   Blab = 'B',
                   DeepShallow = factor('<=100m', levels = c('<=100m', '>100m')))

shallowdeepconstrainedRegression<-ggplot(small, aes(x=log10BACTERIA, y=log10VIRUS)) +
  geom_point(size=.01, alpha=.01) +
  facet_grid(. ~ DeepShallow) +
  geom_abline(data=ann4, aes(intercept=m1int, slope=m1slope), col='#D55E00', size=1) + #orange power law line
  #geom_abline(data=ann4, aes(intercept=MEintset, slope=MEslopeset), col='#33CC33', size=1) + #green mixed effects Fixed Effects line
  geom_abline(intercept=1, slope=1, col='black', size=1) + #black line
   geom_text(data=ann5,aes(x=toX,y=toY,label='10:1'),show_guide=F,size=4,family="Helvetica", fontface="bold") + 
   geom_text(data=ann5,aes(x=plX,y=plY,label='Power Law'),show_guide=F,size=4,family="Helvetica", fontface="bold") +
  geom_text(data=ann5,aes(x=AX,y=AY,label=Alab),show_guide=F,size=4,family="Helvetica", fontface="bold") +
  geom_text(data=ann5,aes(x=BX,y=BY,label=Blab),show_guide=F,size=4,family="Helvetica", fontface="bold") +
  geom_abline(data=ann4, aes(intercept=constrainedints1, slope=m1slope), col='darkgreen', size=.25) +
  geom_abline(data=ann4, aes(intercept=constrainedints2, slope=m1slope), col='darkgreen', size=.25) +
  geom_abline(data=ann4, aes(intercept=constrainedints3, slope=m1slope), col='darkgreen', size=.25) +
  geom_abline(data=ann4, aes(intercept=constrainedints4, slope=m1slope), col='darkgreen', size=.25) +
  geom_abline(data=ann4, aes(intercept=constrainedints5, slope=m1slope), col='darkgreen', size=.25) +
  geom_abline(data=ann4, aes(intercept=constrainedints6, slope=m1slope), col='darkgreen', size=.25) +
  geom_abline(data=ann4, aes(intercept=constrainedints7, slope=m1slope), col='darkgreen', size=.25) +
  geom_abline(data=ann4, aes(intercept=constrainedints8, slope=m1slope), col='darkgreen', size=.25) +
  labs(list(x=expression(paste("Microbe abundance per mL (",log[10]," scale)")),
            y=expression(paste("Virus abundance per mL (",log[10]," scale)")))) +
  theme(aspect.ratio=1,
        title = element_text(size=20),
        strip.background = element_rect(fill="white"),
        strip.text.x = element_text(size = 16,face="bold"),
        axis.text=element_text(size=16, colour="black"),
        axis.title=element_text(size=16,face="bold"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.margin = unit(1, "lines"),
        axis.line = element_line(colour = "black"),
        legend.position = 'none',
        legend.key = element_rect(fill = "white"),
        legend.key.size = unit(.5, "cm")) +
    scale_x_continuous(breaks = c(seq(3.0,8,by=.25)),
                     labels = c(insert_minor(seq(3.0,8,by=.5),1,1)),
                     #limits = c(4.6,8),
                     expand = c(0,0)) +
  scale_y_continuous(breaks = c(seq(3.0,9,by=.25)),
                     labels = c(insert_minor(seq(3.0,9,by=.5),1,1)),
                     #limits = c(4.6,9),
                     expand = c(0,0))
shallowdeepconstrainedRegression

pdf("output/figures/pdf/shallowdeepconstrainedRegression.pdf", width=10, height=5)
print(shallowdeepconstrainedRegression)
dev.off()

jpeg("output/figures/jpg/shallowdeepconstrainedRegression.jpg", width=800, height=400)
print(shallowdeepconstrainedRegression)
dev.off()
```

Comparison of AIC and BIC of each of the linear models applied to viral abundance data
```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SMALLLABSET.RData")
shallowVariableInterceptWithConstraints <- readRDS("output/data/SHALLOWVARIABLEINTERCEPTWITHCONSTRAINTS.rds")
deepVariableInterceptWithConstraints <- readRDS("output/data/DEEPVARIABLEINTERCEPTWITHCONSTRAINTS.rds")

AICc <- function(col1,col2){
  N<-length(col1)
  SSE<-sum(log(col1/col2)^2)
  return(N*log(SSE/N))
}

calcRSQ <- function(obs,pred){
  ssTotal <- sum((obs-mean(obs))^2)
  ssResidual <- sum((obs - pred)^2)
  return(1-(ssResidual/ssTotal))
}

m1<-lm(log10VIRUS~log10BACTERIA,data=smallLabSet)
fm1<-lme(log10VIRUS~log10BACTERIA, random=~1|Study,data=smallLabSet)
fm2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=smallLabSet)

TenOneAICc <-AICc(smallLabSet$log10VIRUS,smallLabSet$log10BACTERIA + 1)
powerlawAICc <-AICc(smallLabSet$log10VIRUS,m1$fitted.values)
VarIntSlopeStudyAICc <- AICc(smallLabSet$log10VIRUS,fm2$fitted[,2])
VarIntFixedSlopeFixedAICc <- AICc(smallLabSet$log10VIRUS,shallowVariableInterceptWithConstraints)

TenOneRSQ <- calcRSQ(smallLabSet$log10VIRUS,smallLabSet$log10BACTERIA + 1)
powerlawRSQ <- calcRSQ(smallLabSet$log10VIRUS,m1$fitted.values)
VarIntSlopeStudyRSQ <- calcRSQ(smallLabSet$log10VIRUS,fm2$fitted[,2])
VarIntFixedSlopeFixedRSQ <- calcRSQ(smallLabSet$log10VIRUS,shallowVariableInterceptWithConstraints)

AIC<-sprintf("%.2f", c(round(TenOneAICc,2),round(powerlawAICc,2),round(VarIntFixedSlopeFixedAICc,2), round(VarIntSlopeStudyAICc,2)))
RSQ<-sprintf("%.2f", c(round(TenOneRSQ,2),round(powerlawRSQ,2),round(VarIntFixedSlopeFixedRSQ,2),round(VarIntSlopeStudyRSQ,2)))
model<-c('10:1','Power Law','Constrained Power Law by Study','Power Law by Study')
modelcomparison<-as.data.frame(cbind(model,RSQ,AIC))
modelcomparison1<-as.data.frame(cbind(model,RSQ,AIC))
kable(modelcomparison1)

load("output/data/DEEPSMALL.RData")
fm1<-lme(log10VIRUS~log10BACTERIA, random=~1|Study,data=DeepSmall)
fm2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=DeepSmall)

TenOneAICc <- AICc(DeepSmall$log10VIRUS,DeepSmall$log10BACTERIA + 1)
powerlawAICc <-AICc(DeepSmall$log10VIRUS,m1$fitted.values)
VarIntSlopeStudyAICc <-AICc(DeepSmall$log10VIRUS,fm2$fitted[,2])
VarIntFixedSlopeFixedAICc <-AICc(DeepSmall$log10VIRUS,deepVariableInterceptWithConstraints)

TenOneRSQ <- calcRSQ(DeepSmall$log10VIRUS,DeepSmall$log10BACTERIA + 1)
powerlawRSQ <- calcRSQ(DeepSmall$log10VIRUS,m1$fitted.values)
VarIntSlopeStudyRSQ <- calcRSQ(DeepSmall$log10VIRUS,fm2$fitted[,2])
VarIntFixedSlopeFixedRSQ <- calcRSQ(DeepSmall$log10VIRUS,deepVariableInterceptWithConstraints)

AIC<-sprintf("%.2f", c(round(TenOneAICc,2),round(powerlawAICc,2),round(VarIntFixedSlopeFixedAICc,2), round(VarIntSlopeStudyAICc,2)))
RSQ<-sprintf("%.2f", c(round(TenOneRSQ,2),round(powerlawRSQ,2),round(VarIntFixedSlopeFixedRSQ,2),round(VarIntSlopeStudyRSQ,2)))
model<-c('10:1','Power Law','Constrained Power Law by Study','Power Law by Study')
modelcomparison<-as.data.frame(cbind(model,RSQ,AIC))
modelcomparison2<-as.data.frame(cbind(model,RSQ,AIC))
kable(modelcomparison2)

names(modelcomparison1) <- c("Model", "Shallow RSQ", "Shallow AIC")
names(modelcomparison2) <- c("Model", "Deep RSQ", "Deep AIC")
totalAIC<-merge(modelcomparison1,modelcomparison2, all = TRUE)
kable(totalAIC[c(1,3,2,4),])
```


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
#figure31
load("output/data/SMALLLABSET.RData")
load("output/data/DEEPSMALL.RData")
fm2<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=smallLabSet)
fm2a<-lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=DeepSmall)
s2<-summary(fm2)
s2f<-as.data.frame(s2$coefficients$fixed)
s2r<-as.data.frame(s2$coefficients$random)
names(s2f) <- c("commonADJ")
names(s2r) <- c("Shallow Intercept", "Shallow Slope")
s2full<-as.data.frame(cbind(rownames(s2r),sprintf("%.2f",s2r$'Shallow Intercept' + s2f$commonADJ[1]), sprintf("%.2f",s2r$'Shallow Slope' + s2f$commonADJ[2])))
names(s2full) <- c("Study", "Shallow Intercept", "Shallow Slope")
s2a<-summary(fm2a)
s2fa<-as.data.frame(s2a$coefficients$fixed)
s2ra<-as.data.frame(s2a$coefficients$random)
names(s2fa) <- c("commonADJ")
names(s2ra) <- c("Shallow Intercept", "Shallow Slope")
s2fulla<-as.data.frame(cbind(rownames(s2ra),sprintf("%.2f",s2ra$'Shallow Intercept' + s2fa$commonADJ[1]), sprintf("%.2f",s2ra$'Shallow Slope' + s2fa$commonADJ[2])))
names(s2fulla) <- c("Study", "Deep Intercept", "Deep Slope")

totalcv<-merge(s2full,s2fulla, all = TRUE)
totalcv[is.na(totalcv)] <- NA
totalcv1 <- sapply(totalcv, as.character)
totalcv1[is.na(totalcv1)] <- " "
kable(totalcv1)
```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/DEEPCONSTRAINTS.RData")
load("output/data/SHALLOWCONSTRAINTS.RData")

shallowcv<-shallowvariableInterceptwithconstraints
deepcv<-deepvariableInterceptwithconstraints
shallowtab<-as.data.frame(summary(shallowcv)$parameters)
shallowtab<-as.data.frame(cbind(Study=rownames(shallowtab),sprintf("%.2f",shallowtab[,1]),sprintf("%.2f",shallowtab[,2])))
deeptab<-as.data.frame(summary(deepcv)$parameters)
deeptab<-as.data.frame(cbind(Study=rownames(deeptab),sprintf("%.2f",deeptab[,1]),sprintf("%.2f",deeptab[,2])))
names(shallowtab) <- c("Study", "Shallow Est", "Shallow StErr")
names(deeptab) <- c("Study", "Deep Est", "Deep StErr")
totalcv<-merge(shallowtab,deeptab, all = TRUE)
totalcv[is.na(totalcv)] <- NA
totalcv1 <- sapply(totalcv, as.character)
totalcv1[is.na(totalcv1)] <- " "
kable(totalcv1)

```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load('output/data/SHALLOWRSQFITS.RData')
load('output/data/DEEPRSQFITS.RData')

shallowrsqfits<-rsqfits
names(shallowrsqfits) <- c("Study", "Shallow RSQ", "Shallow P-Value")
#deepobs<-as.data.frame(table(DeepSmall$Study))
names(deeprsqfits) <- c("Study", "Deep RSQ", "Deep P-Value")
totalrsqs<-merge(shallowrsqfits,deeprsqfits, all = TRUE)
totalrsqs[is.na(totalrsqs)] <- NA
totalrsqs1 <- sapply(totalrsqs, as.character)
totalrsqs1[is.na(totalrsqs1)] <- " "
kable(totalrsqs1)
```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/SHALLOWPERCENTILES2.RData")
load("output/data/DEEPPERCENTILES2.RData")
#percentiles2
#deeppercentiles2
ninetfyfivepercent<-merge(percentiles2,deeppercentiles2, all = TRUE, by="row.names")
kable(ninetfyfivepercent)
```

```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
load("output/data/DEEPCONSTRAINTS.RData")
load("output/data/SHALLOWCONSTRAINTS.RData")
load("output/data/SMALLLABSET.RData")
deepvariableInterceptwithconstraints

summobj<-summary(shallowvariableInterceptwithconstraints)
df<-as.data.frame(summary(shallowvariableInterceptwithconstraints)$coefficients, rownames=F)
s<-df[,1:2]
s <- as.data.frame(cbind(Study = rownames(s), s))
s$Study <- factor(s$Study, levels = c("slope", "ARCTICSBI","BATS","BEDFORDBASIN","CASES0304","ELA","FECYCLE1","FECYCLE2","GEOTRACES","GEOTRACES_LEG3","GREENLAND2012","INDIANOCEAN2006","KHO4","KH05_2","MOVE","NASB2005","NORTHSEA2001","POWOW","Raunefjord","SOG","STRATAPHYT1","STRATAPHYT2","SWAT","TABASCO","TROUT","USCMO"))
s<-s[order(s$Study),]
rownames(s) <- NULL
#kable(s)
m1<-lm(log10VIRUS~log10BACTERIA,data=smallLabSet)
s1<-s[2:nrow(s),]
s1<-s1[rev(order(s1$Estimate)),]
m1row <- data.frame(Study = 'Power Law REGRESSION INT', Estimate = coef(m1)[[1]], 'Std. Error' = summary(m1)$coefficients[ , 2][1], check.names=F)
s1<-rbind(s1,m1row)
#rsqtable<-rsqfits[order(as.numeric(format(pvals, scientific = TRUE, digits = 3))),]
s1<-s1[rev(order(s1$Estimate)),]
rownames(s1) <- NULL
s1<-cbind(s1,Group=c(rep("A",9)," ",rep("B",16)))
kable(s1)
```


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}
#identify inliers
#table(small$DepthCategory)
load("output/data/SMALL.RDATA")
load("output/data/SMALLLABSET.RDATA")
load("output/data/DEEPSMALL.RDATA")
low<-quantile(small$Log10_VirBac_Ratio,probs=c(.02))
high<-quantile(small$Log10_VirBac_Ratio,probs=c(.98))
inset<-smallLabSet[smallLabSet$Log10_VirBac_Ratio <= high & smallLabSet$Log10_VirBac_Ratio >= low,]
small$inset<-ifelse(small$Log10_VirBac_Ratio <= high & small$Log10_VirBac_Ratio >= low,'Inlier Set','Outlier Set')
insetDeep<-small[which(small$Log10_VirBac_Ratio <= high & small$Log10_VirBac_Ratio >= low & small$DepthCategory == ">100m"),]
smallInset<-small[which(small$Log10_VirBac_Ratio <= high & small$Log10_VirBac_Ratio >= low),]

ann4 <- data.frame(
  intset = c(summary(lm(inset$log10VIRUS ~ inset$log10BACTERIA))$coefficients[1],summary(lm(insetDeep$log10VIRUS ~ insetDeep$log10BACTERIA))$coefficients[1]),
  slopeset = c(summary(lm(inset$log10VIRUS ~ inset$log10BACTERIA))$coefficients[2],summary(lm(insetDeep$log10VIRUS ~ insetDeep$log10BACTERIA))$coefficients[2]),
  ALLintset = c(summary(lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA))$coefficients[1],summary(lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA))$coefficients[1]),
  ALLslopeset = c(summary(lm(smallLabSet$log10VIRUS ~ smallLabSet$log10BACTERIA))$coefficients[2],summary(lm(DeepSmall$log10VIRUS ~ DeepSmall$log10BACTERIA))$coefficients[2]),
  MEintset= c(summary(lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=inset))$coefficients$fixed[1],summary(lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=insetDeep))$coefficients$fixed[1]),
  MEslopeset = c(summary(lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=inset))$coefficients$fixed[2],summary(lme(log10VIRUS~log10BACTERIA, random=~1+log10BACTERIA|Study,data=insetDeep))$coefficients$fixed[2]),
  DeepShallow = c('<=100m', '>100m'),
  BacMedians = c(median(inset$log10BACTERIA), median(insetDeep$log10BACTERIA)),
  VirMedians = c(median(inset$log10VIRUS), median(insetDeep$log10VIRUS))
  )

ann5 <- data.frame(meX = 3.73, 
                   meY = 5.3, 
                   plX = 6.8,
                   plY = 6.9, 
                   toX=7, 
                   toY = 8.3, 
                   DeepShallow = factor('<=100m', levels = c('<=100m', '>100m')))
smallInset$DeepShallow <- smallInset$DepthCategory
summary(smallInset)
shallowDeepInliers<-ggplot(smallInset, aes(x=log10BACTERIA, y=log10VIRUS)) +
  geom_point(col='steel blue',pch=20, size=1) +
  facet_grid(. ~ DeepShallow) +
  geom_abline(data=ann4, aes(intercept=ALLintset, slope=ALLslopeset), col='#D55E00', size=1) + #orange all data power law line
  geom_abline(data=ann4, aes(intercept=intset, slope=slopeset), col='chartreuse3', size=1) + #green inset power law line
  #geom_abline(data=ann4, aes(intercept=MEintset, slope=MEslopeset), col='#33CC33', size=1) + #green mixed effects Fixed Effects line
  geom_abline(intercept=1, slope=1, col='black', size=1) + #black line
  geom_segment(data=ann4,aes(x=BacMedians,xend=BacMedians,y=4.8,yend=4.9),arrow=arrow(length=unit(0.3,"cm")),show_guide=F, color="black") +
  geom_segment(data=ann4,aes(x=3.5,xend=3.6,y=VirMedians,yend=VirMedians),arrow=arrow(length=unit(0.3,"cm")),show_guide=F, color="black") +
   #geom_text(data=ann5,aes(x=meX,y=meY,label='ME'),show_guide=F,size=6,family="Helvetica", fontface="bold") + 
   geom_text(data=ann5,aes(x=toX,y=toY,label='10:1'),show_guide=F,size=6,family="Helvetica", fontface="bold") + 
   geom_text(data=ann5,aes(x=plX,y=plY,label='Power Law'),show_guide=F,size=6,family="Helvetica", fontface="bold") + 
  labs(list(x=expression('Microbes per mL (log'[10]*' scale)'), 
            y=expression('Viruses per mL (log'[10]*' scale)')))+
  theme(panel.margin = unit(2, "lines"),
        strip.background = element_rect(fill="white"),
        strip.text.x = element_text(size = 20,face="bold"),
          aspect.ratio=1,
          plot.title=element_text(size=32),
          axis.text=element_text(size=12), 
          axis.title=element_text(size=16,face="bold"),
          #text = element_text(size=32),aspect.ratio=1,
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black")) +
    scale_x_continuous(breaks = c(seq(2.5,8,by=.25)), 
                     labels = c(insert_minor(seq(2.5,8,by=.5),1,1)), 
                     limits = c(3.5,7.5), 
                     expand = c(0,0)) +
    scale_y_continuous(breaks = c(seq(4.5,9,by=.25)), 
                     labels = c(insert_minor(seq(4.5,9,by=.5),1,1)), 
                     limits = c(4.8,8.75), 
                     expand = c(0,0))
shallowDeepInliers
pdf("output/figures/pdf/shallowDeepInliers.pdf", width=10, height=7)
print(shallowDeepInliers)
dev.off()

jpeg("output/figures/jpg/shallowDeepInliers.jpg", width=800, height=560)
print(shallowDeepInliers)
dev.off()

kable(ann4)

#write.table(data, "VIRBACDATA.txt", sep=",")
```


```{r, echo = FALSE, warning=F, results='hold', message=FALSE}

load("output/data/DATA.RData")
# plot(sort(data$DEPTH))
# plot(ecdf(data$DEPTH))
# median(data$DEPTH)
# 
# levels(data$Study)
# require(plyr)
# data$Study<-revalue(data$Study, c("Balsom - arctic"="ARCTICSBI",'Bratbak 2001'='NORTHSEA2001', "MiddleboeGreenland2012"="GREENLAND2012", "MiddleBoeIndia" = 'INDIANOCEAN2006', "BENFORDBASIN" = 'BEDFORDBASIN', "FeCycle   "= "FECYCLE1", "FeCycle II"= "FECYCLE2","Raunefjord 2000"="RAUNEFJORD2000","Stratiphyt I"="STRATIPHYT1","Stratiphyt II"="STRATIPHYT2","Trout"="TROUT","GeoTraces"="GEOTRACES","GeoTraces_Leg3"="GEOTRACES_LEG3"))
# index <- which(data$Study == 'WINTER - DYFAMED' | data$Study == 'DFO')
# data <- data[-index, ]
# data<-data[which(!is.na(data$BACTERIA) & !is.na(data$VIRUS)), ]
# data$Study<-droplevels(data$Study)
pdf("output/figures/pdf/DepthMedianPlot.pdf", width=6, height=6)
plot(sort(log10(data$DEPTH)), pch=20, ylab='Log10 Depth', col='steelblue')
abline(h=median(log10(data$DEPTH)), col='black')
abline(v=length(data$DEPTH)/2, col='black')
dev.off()
# plot(ecdf(log10(data$DEPTH)), xlab='Log10 Depth', main='')
# abline(h=.5, col='chartreuse3')
# abline(v=median(log10(data$DEPTH)), col='chartreuse3')

data$log10BACTERIA<-log10(data$BACTERIA)
data$log10VIRUS<-log10(data$VIRUS)
lm(data$log10VIRUS~data$log10BACTERIA)
plot(data$log10VIRUS~ data$log10BACTERIA, pch=46, col='steelblue')


shallowslopes<-c()
deepslopes<-c()
cutoff<-seq(50,150)
cutoffquantiles<-c()
for (i in cutoff){
  s<-summary(lm(data$log10VIRUS[which(data$DEPTH<=i)]~data$log10BACTERIA[which(data$DEPTH<=i)]))
  shallowslope<-s$coefficients[2]
  shallowslopes<-c(shallowslopes,shallowslope)

  s<-summary(lm(data$log10VIRUS[which(data$DEPTH>i)]~data$log10BACTERIA[which(data$DEPTH>i)]))
  deepslope<-s$coefficients[2]
  deepslopes<-c(deepslopes,deepslope)
  cutoffquantiles<-c(cutoffquantiles,length(data$log10VIRUS[which(data$DEPTH<=i)])/length(data$log10VIRUS))
}
cbind(shallowslopes,deepslopes,cutoff)
pdf("output/figures/pdf/SlopesByCutoff.pdf", width=6, height=6)
plot(cutoff,shallowslopes, pch=20, col='steelblue', xlim = c(50,150), ylim = c(0,1), axes=T,
     xlab='Depth Threshold (m)', ylab='Power-law exponent')
points(cutoff,deepslopes, pch=20, col='chartreuse3')
axis(side=1, at=c(seq(50,150,length=6)))
axis(side=2, at=c(seq(0,1,length=6)))
legend("bottomleft", inset=.05,
       c("Shallow","Deep"), fill=c('steelblue','chartreuse3'), horiz=F, bty = "n")
dev.off()

pdf("output/figures/pdf/SlopesByCutoffALT.pdf", width=6, height=6)
#par(bty="l")
plot(cutoff,shallowslopes, pch=20, col='steelblue', ylim = c(0,1), axes=T,
     xlab='Depth Threshold (m)', ylab='Power-law exponent', bty="l")
#axis(side=1, at=c(seq(50,150,length=6)))
#axis(side=2, at=c(seq(0,1,length=6)))
axis(1, xaxp = c(50, 150, 5))
axis(2, xaxp = c(0, 1, 9))
points(cutoff,deepslopes, pch=20, col='chartreuse3')
#legend("topright", inset=.05, title="Slopes by cutoff point",
legend("bottomleft", inset=.05,
       c("Shallow","Deep"), fill=c('steelblue','chartreuse3'), horiz=F, bty = "n")
dev.off()
```
